---
title: "Final figures paper: Genotype-phenotype single-cell transcriptomics for massive parallel assessment of genetic variants "
output: html_document
date: "2026-05-04"
author: "Rosa De Santis"
---

#library needed to run the code
```{r setup, include=FALSE}
library(ggplot2)
library(pheatmap)
library(matrixStats)
library(scales)
library(qdap)
library(stringr)
library(dplyr)
library(tidyverse)
library(ggridges)
library(ggplot2)
library(scales)
library(cowplot)
library(dplyr)
library(Seurat)
library(Matrix)
library(reshape2)
library(edgeR)
library(matrixStats)
library(readr)
library(enrichR)
library(tibble)
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(ReactomePA)
library(msigdbr)
library(edgeR)
library(dplyr)
library(Seurat)
library(Matrix)
library(reshape2)
library(edgeR)
library(matrixStats)
library(tidyr)
library(Biostrings)
library(scuttle)
library(SingleCellExperiment)
library(ggnewscale)
library(patchwork)
```

To reproduce the figures from GEO data, the following steps are required:

Download raw FASTQ files from GEO 
Run alignment and variant calling using the SCRAMseq pipeline (described in Methods; scripts available at github), which produces for each experiment:

{domain}_mut_list_muscan_single_rep.txt — MITE enrichment scores (logFC per variant, edgeR output)
{domain}_BC_mut_assignment_with_effect.rds — cell barcode to variant assignment with functional classification
mutations_{experiment}_to_conversion.txt — per-cell SCRAMseq scores (Sign_test_11) with variant assignment

The other files are available as supplemetary tables

Figure 2E top - CD molecules expression
```{r}
fig="1E_top"

res=read.table("Results/WTvsGFP_results_annotated.txt",sep="\t",header=T)
cd_ann=read.table("Data/CD_proteins_annotation.txt",header=T,quote="",sep="\t")
res$CD=ifelse(res$external_gene_name%in%cd_ann$Approved.symbol,T,F)

res=res[order(res$logFC_WTvsGFP,decreasing=T),]
res_sub=res[res$CD,][1:5,]

plot=cbind(melt(res_sub[,2:4],id.vars = "external_gene_name"),sd=c(res_sub$Sd_WT,res_sub$Sd_GFP))
plot$variable=gsub("Avg_","",plot$variable)
plot$external_gene_name=factor(plot$external_gene_name,levels=res_sub$external_gene_name)

colors=c("blue4","red4")

pdf(paste0("Figure1/",fig,".pdf"),width=6,height=6)
ggplot(plot, aes(x=external_gene_name, y=value, fill=variable))+
    geom_bar(color="black", position = position_dodge(),stat = "identity",width = .5)+
  theme_classic()+
  theme(axis.title = element_blank(),
        legend.position = "none")+
  geom_errorbar(aes(ymin=value, ymax=value+sd), width=.2, position = position_dodge(.5))+
  scale_fill_manual(values=colors)
dev.off()
```

Fig. 2F FACS
```{r}

fig="1E_btm"

data.path = "Data/FACS_wt_gfp"
fs <- read.flowSet(path = data.path, transformation = "linearize",  
                   emptyValue = FALSE,  alter.names = TRUE,  
                   column.pattern = ".H|FSC|SSC") 
sampleNames(fs)=gsub(" ","_",sampleNames(fs))

p=list()
for(i in 1:length(fs)){
  data=as.data.frame(fs[[i]]@exprs)
  data$sample=sapply(strsplit(sampleNames(fs)[i], "_"), "[[", 3)
  p[[i]]=data
}

final=multi_join(p, rbind)

final_sub=final[final$sample%in%c("WT","EMPTY"),]
threshold=quantile(final_sub[final_sub$sample=="WT","PE.H"],.25)

colors=c(alpha("blue4",.3),alpha("red4",.3))
  
pdf(paste0("Figure1/",fig,".pdf"),width=16,height=6)
  ggplot(final_sub, aes_string(x="PE.H", fill="sample",color="sample"))+geom_histogram(bins = 300,alpha=.3)+
    scale_x_continuous(trans = "log10")+
    theme_classic()+theme(axis.title = element_blank())+
    geom_vline(xintercept = threshold)+
    scale_fill_manual(values=colors)+
    scale_color_manual(values=colors)+
    facet_wrap(~sample)
  dev.off()

```



Fig. 2D Volcano Plot and S1A GO enrichment
```{r}
if(!file.exists("Results/WTvsGFP_results_annotated.txt")){
mtx=read.table("Paper_FP/Data/P63_Mutants_validation_final_merged_rawCountsWithAnnotations.txt",header=T,sep="\t",quote="",stringsAsFactors = F)

pjsamples=data.frame(unique_samples=colnames(mtx)[-c(1:2)], condition=ifelse(stringr::str_count(colnames(mtx)[-c(1:2)],"_")>1,sapply(strsplit(colnames(mtx)[-c(1,2)],"_"),"[[", 2),sapply(strsplit(colnames(mtx)[-c(1,2)],"_"),"[[", 1)))

mtx_red=mtx[,c(1,2,grep("WTsec|WTthir|GFP",colnames(mtx)))]
pjsamples_red=pjsamples[pjsamples$unique_samples%in%colnames(mtx_red),]
pjsamples_red$condition=ifelse(grepl("WT",pjsamples_red$condition),"WT","GFP")
cpm=cbind(mtx_red[,1:2], cpm(mtx_red[,-c(1:2)]))
## Average values
avg=as.data.frame(sapply(unique(pjsamples_red$condition), function(x) rowMeans(cpm[colnames(cpm)%in%pjsamples_red[pjsamples_red$condition==x,1]])))
colnames(avg)=paste0("Avg_",colnames(avg))
sd=as.data.frame(sapply(unique(pjsamples_red$condition), function(x) rowSds(as.matrix(cpm[colnames(cpm)%in%pjsamples_red[pjsamples_red$condition==x,1]]))))
colnames(sd)=paste0("Sd_",colnames(sd))

ann=cbind(cpm[,1:2],avg,sd)

wt="GFP"
data=mtx_red[,-c(1:2)]
rownames(data)=mtx_red[,1]
res=de_unpaired(data,pjsamples_red$condition,ann=ann)
write.table(res,"Paper_FP/Results/WTvsGFP_results_annotated.txt",row.names = F, sep="\t",quote=F)
} else {
  res=read.table("Paper_FP/Results/WTvsGFP_results_annotated.txt",sep="\t",header=T)
}

fig="1D"

condition="WTvsGFP"
volc=res[,c(2,grep(condition, colnames(res)))]
volc=volc[complete.cases(volc),]
down=sum(volc[,4]=="DOWN")
up=sum(volc[,4]=="UP")

volc[,3][volc[,3]==0]= min(volc[,3][volc[,3]!=0 & !is.na(volc[,3])])

volc=volc[order(volc[,2],decreasing=T),]
volc[1:10,4]="UP-Top"
volc[(nrow(volc)-10):nrow(volc),4]="DOWN-Top"

pdf(paste0("Figure1/",fig,".pdf"),height=6, width=8)
  ggplot(volc, aes(x = volc[,2], y = -log10(volc[,3]), color = volc[,4], size=volc[,4]))+  
  geom_point(na.rm = T) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none")+
  xlab(paste0("LFC (", condition, ")")) +
  ylab(expression(-log[10]("FDR"))) +
  geom_hline(yintercept = 1.3, colour="#990000", linetype="dashed") + geom_vline(xintercept = 1.5, colour="#990000", linetype="dashed") + geom_vline(xintercept = -1.5, colour="#990000", linetype="dashed") +
  scale_color_manual(values = c("grey70", "blue4", "blue4","red4","red4")) +
    scale_size_manual(values = c(1,1,1,1,3))+
  scale_y_continuous(trans = "log1p") +
    annotate(geom="text", x=min(volc[,2],na.rm=T)/1.2, y=1.6, label=as.character(down,na.rm=T), color="blue4", size=8)+
  annotate(geom="text", x=max(volc[,2],na.rm=T)/1.2, y=1.6, label=as.character(up, na.rm=T), color="red4",size=8)
  dev.off()
  
list_genes=volc[grep("UP",volc$diffEXP_WTvsGFP),1]

dbs <- c("GO_Biological_Process_2023")
  enriched <- enrichr(list_genes, dbs)
  
  p=list()
  for(i in 1:length(enriched)){
  enriched[[i]]$count=as.numeric(sapply(strsplit(as.character(enriched[[i]]$Overlap),"/"), "[[", 1))
  enriched[[i]]$total=as.numeric(sapply(strsplit(as.character(enriched[[i]]$Overlap),"/"), "[[", 2))
  enriched[[i]]$perc=enriched[[i]]$count/enriched[[i]]$total*100
  test=enriched[[i]]
  test=test[order(test$Adjusted.P.value, decreasing = F),]
  test$Term=factor(test$Term,levels=rev(test$Term))
  test$db=dbs[i]
  p[[i]]=test
  }
  kegg=p[[1]]
  kegg=kegg[order(kegg$Adjusted.P.value, decreasing = F),]
  kegg=kegg[c(4,5,10,11,14),c(1,3,12,4,13)]
  colnames(kegg)=c("Path","p.value","overlap","p.adj","db")
  
  data=kegg
  data$Path=factor(data$Path,levels=rev(data$Path))

plot=ggplot(data, aes(x = -log10(p.adj), y = Path, fill=overlap)) +
  theme_bw()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size=10),
        strip.background = element_blank(),
  strip.text = element_blank()) +
  ylab("") +
  xlab("")+
  labs(color = "")+
  geom_bar(color="black",stat = "identity",width=.7) +
  scale_fill_distiller(type = "seq", palette = "Reds",direction = 1)+
  geom_vline(xintercept = -log10(0.05),linetype="dashed",color="black")

 pdf(paste0("Figure1/",fig, "_GO.pdf"), height=4, width=8)
  print(plot)
  dev.off()

fig="S1A"

kegg=p[[1]]
  kegg=kegg[order(kegg$Adjusted.P.value, decreasing = F),]
  kegg=kegg[kegg$Adjusted.P.value<0.05,c(1,3,12,4,13)]
  colnames(kegg)=c("Path","p.value","overlap","p.adj","db")
  
  data=kegg
  data$Path=sapply(strsplit(as.character(data$Path),"[(]"),"[[",1)
  data$Path=gsub("[)]","",data$Path)
  data$Path=factor(data$Path,levels=rev(data$Path))

plot=ggplot(data, aes(x = -log10(p.adj), y = Path, fill=overlap, label=Path)) +
  theme_bw()+
  theme(axis.title = element_blank(),
        axis.text =  element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size=10),
        strip.background = element_blank(),
  strip.text = element_blank()) +
  ylab("") +
  xlab("")+
  labs(color = "")+
  geom_bar(color="black",stat = "identity",width=.7) +
  scale_fill_distiller(type = "seq", palette = "Reds",direction = 1)+
  geom_text()+
  geom_vline(xintercept = -log10(0.05),linetype="dashed",color="black")

 pdf(paste0("Figure1/",fig, "_GO.pdf"), height=12, width=6)
  print(plot)
  dev.off()

```


Figure S3A heatmap SAM/DBD and Figure S3B scatter correlation 
```{r}
fig="S2A"
domain="SAM"

#Phenotypic score
#input used for this plot can be obtained from the processing of MITE experiment
clust=read.table(paste0("Data/",domain,"_mut_list_muscan_single_rep.txt"),header = T)

sub=clust[,"logFC",drop=F]
sub=sub[rownames(sub)!="WT",,drop=F]
sub$residues=substr(rownames(sub), 1, 4)
sub$aa=substr(rownames(sub), 5, 5)
sub$order=as.numeric(substr(sub$residues, 2,4))+43
sub$aa_or=substr(sub$residues, 1,1)
rownames(sub)=paste0(sub$aa_or,sub$order,sub$aa)

#Predictors scores
input=read.table(paste0("Predictors_benchmark_RDS/Mutpred_prediction",domain,".txt"), row.names = 1,header=T)
input=pred[,-c(7,8)]

sort=read.table("Results/order_aa_heatmaps.txt",header=T)
sort=as.character(sort$aa)

p1=list()

p2=list()
for(i in 1:8){
mut_pred=input[,i,drop=F]
mut_pred$residues=substr(rownames(mut_pred), 1, 4)
mut_pred$aa=substr(rownames(mut_pred), 5, 5)
mut_pred$order=substr(mut_pred$residues, 2,4)
mut_pred=mut_pred[order(mut_pred$order, decreasing=T),]
colnames(mut_pred)[1]="score"
mut_pred$residues=factor(mut_pred$residues, levels = rev(unique(mut_pred$residues)))


mut_pred$aa=factor(mut_pred$aa, levels = rev(sort))

corr=merge(mut_pred, sub[,"logFC",drop=F],by=0)
print(paste0("Pearson correlation is ", cor(as.numeric(corr$score), corr$logFC, method = "pearson", use="pairwise.complete.obs"), " for ",colnames(input)[i]))

z <- as.data.frame(complete(mut_pred, residues, aa))

if(i==1){
limit=0.8
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)
} else if(i==2) {
#min=-6
limit=(-2.2)
colors=colorRampPalette(c("red2", 'red2', 'white', "darkblue", "darkblue"))(100)
} else if(i==3) {
limit=0.05
colors=colorRampPalette(c("red2", 'red2', 'white', "darkblue", "darkblue"))(100)
} else if(i==7) {
limit=20
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)
} else if (i==8) {
limit=(-7.5)
colors=colorRampPalette(rev(c("darkblue", "darkblue", 'white', "red2", 'red2')))(100)
}else {
limit=0.5
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)
}


library(scales)
  if (i == 2) {
    #z$score[z$score < -6] <- -6
    p1[[i]] <- ggplot(z, aes(aa, residues)) +
      theme_void() +
      geom_tile(aes(fill = as.numeric(score)), size = 1, color = "white") +
      scale_fill_gradientn(
        colors = colors,
        values = rescale(c(-12, limit, max(as.numeric(z$score[!is.na(z$score)])))),
        limits = c(-12, max(as.numeric(z$score[!is.na(z$score)]))),  
        labels=scales::label_number(accuracy = 0.1), #0.1 for grid arrange
        na.value = "grey50"
      ) +
      ylab("") +
      xlab("") +
      theme(
        axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 15, family = "Courier"),
        legend.title = element_blank(),
        legend.key.size = unit(.5, "cm"),
        legend.justification = c("right", "bottom"),
        legend.text = element_text(size = 8, family = "Courier") #zsize 8 for grid arrange, instead 13
      ) +
      coord_flip()
       p2[[i]] <- ggplot(corr, aes(x = logFC, y = as.numeric(score))) +
      geom_point(size = 0.5, color = "black") +
      theme_bw() +
      theme(
        panel.grid = element_blank(),
        axis.text = element_text(size = 9),
        axis.title = element_text(size = 15)
      ) +
      xlab("Phenotypic score") +
      ylab("Predictor score") +
      ggtitle(colnames(input)[i]) +
      geom_vline(xintercept = 0, linetype = "dashed", color = "black", size = 0.5) +
      geom_hline(yintercept = limit, linetype = "dashed", color = "black", size = 0.5) +
      scale_x_continuous(limits = c(-5, 6))+
               scale_y_continuous(limits = c(-12, 1), 
                                  breaks = seq(-12, 1, by = 3),expand = c(0,0))+
         theme(plot.margin = margin(5, 5, 5, 5))

  } else if (i == 7) {
    p1[[i]] <- ggplot(z, aes(aa, residues)) +
      theme_void() +
      geom_tile(aes(fill = as.numeric(score)), size = 1, color = "white") +
       scale_fill_gradientn(
    colors = colors,
    values = scales::rescale(c(0, limit, max(as.numeric(z$score), na.rm = TRUE))),
    limits = c(0, max(as.numeric(z$score), na.rm = TRUE)),
    labels = scales::label_number(accuracy = 0.1),
    na.value = "grey50"
  )+
      ylab("") +
      xlab("") +
      theme(
        axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 15, family = "Courier"),
        legend.title = element_blank(),
        legend.key.size = unit(.5, "cm"),
        legend.justification = c("right", "bottom"),
        legend.text = element_text(size = 15, family = "Courier")
      ) +
      coord_flip()
    

    p2[[i]] <- ggplot(corr, aes(x = logFC, y = as.numeric(score))) +
      geom_point(size = 0.5, color = "black") +
      theme_bw() +
      theme(
        panel.grid = element_blank(),
        axis.text = element_text(size = 9.5),
        axis.title = element_text(size = 15)
      ) +
      xlab("Phenotypic score") +
      ylab("Predictor score") +
      ggtitle(colnames(input)[i]) +
      geom_vline(xintercept = 0, linetype = "dashed", color = "black", size = 0.5) +
      geom_hline(yintercept = limit, linetype = "dashed", color = "black", size = 0.5) +
      scale_x_continuous(limits = c(-5, 6)) +
      scale_y_continuous(limits = c(0, 30), labels=scales::label_number(accuracy = 0.1)) +
      theme(plot.margin = margin(5, 5, 5, 5))
  } else if (i == 8) {
    p1[[i]] <- ggplot(z, aes(aa, residues)) +
      theme_void() +
      geom_tile(aes(fill = as.numeric(score)), size = 1, color = "white") +
      scale_fill_gradientn(
        colors = colors,
          values = rescale(c(min(as.numeric(z$score[!is.na(z$score)])), limit, max(as.numeric(z$score[!is.na(z$score)])))),
        limits = c(min(as.numeric(z$score[!is.na(z$score)])), max(as.numeric(z$score[!is.na(z$score)]))),  labels = scales::label_number(accuracy = 1)
      ) +
      ylab("") +
      xlab("") +
      theme(
        axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 15, family = "Courier"),
        legend.title = element_blank(),
        legend.key.size = unit(.5, "cm"),
        legend.justification = c("right", "bottom"),
        legend.text = element_text(size = 18, family = "Courier")
      ) +
      coord_flip()

    p2[[i]] <- ggplot(corr, aes(x = logFC, y = as.numeric(score))) +
      geom_point(size = 0.5, color = "black") +
      theme_bw() +
      theme(
        panel.grid = element_blank(),
        axis.text = element_text(size = 9.5),
        axis.title = element_text(size = 15)
      ) +
      xlab("Phenotypic score") +
      ylab("Predictor score") +
      ggtitle(colnames(input)[i]) +
      geom_vline(xintercept = 0, linetype = "dashed", color = "black", size = 0.5) +
      geom_hline(yintercept = limit, linetype = "dashed", color = "black", size = 0.5) +
      scale_x_continuous(limits = c(-5, 6)) +
      scale_y_continuous(limits = c(-20, 10), labels=scales::label_number(accuracy = 0.1), breaks = seq(-20, 10, by = 5), expand = c(0,0)) +
      theme(plot.margin = margin(5, 5, 5, 5))
    
  } else {
    p1[[i]] <- ggplot(z, aes(aa, residues)) +
      theme_void() +
      geom_tile(aes(fill = as.numeric(score)), size = 1, color = "white") +
      scale_fill_gradientn(
        colors = colors,
        values = rescale(c(min(as.numeric(z$score[!is.na(z$score)])), limit, max(as.numeric(z$score[!is.na(z$score)])))),
         limits = c(0, max(as.numeric(z$score[!is.na(z$score)]))),
        na.value = "grey50"
      ) +
      ylab("") +
      xlab("") +
      theme(
        axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 15, family = "Courier"),
        legend.title = element_blank(),
        legend.key.size = unit(.5, "cm"),
        legend.justification = c("right", "bottom"),
        legend.text = element_text(size = 15, family = "Courier")
      ) +
      coord_flip()

    p2[[i]] <- ggplot(corr, aes(x = logFC, y = as.numeric(score))) +
      geom_point(size = 0.5, color = "black") +
      theme_bw() +
      theme(
        panel.grid = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 15)
      ) +
      xlab("Phenotypic score") +
      ylab("Predictor score") +
      ggtitle(colnames(input)[i]) +
      geom_vline(xintercept = 0, linetype = "dashed", color = "black", size = 0.5) +
      geom_hline(yintercept = limit, linetype = "dashed", color = "black", size = 0.5) +
      scale_x_continuous(limits = c(-5, 6))+
      scale_y_continuous(limits = c(0, 1)) +
      theme(plot.margin = margin(5, 5, 5, 5))
  }
}


pdf(paste0("figures/",fig,"_",domain,"samescales_test.pdf"), height =4,width = 16)
for(i in 1:length(p2)){
  if (i==2){
        print(grid.arrange(p1[[i]],nullGrob(),p2[[i]] , nrow=1, ncol=3, widths=c(2,0.05,1)))
  } else {
    print(grid.arrange(p1[[i]],p2[[i]] , nrow=1, ncol=2, widths=c(2,1)))
  }
}
dev.off()

pdf(paste0("figures/",fig,"_",domain,"corr.pdf"), height =4,width =5 )
for(i in 1:length(p2)){
    print(p2[[i]])
}
dev.off()
```


Fig. S3C Spearman correlation MITE vs all predictors
```{r}
domain="SAM" #or DBD
pred=read.table(paste0("Predictors_benchmark_RDS/",domain,"_all_predictors.txt"), header = T, sep="\t") 
pred$mut_name=rownames(pred)

#DBD sheet 1 SAM sheet 2
eve=read_excel(path = "Output MITE and predictors.xlsx", sheet = 2)
eve$mut_name=eve$Variant

pred_joined=pred %>%
  left_join(eve %>% dplyr::select(mut_name, EVE), by = "mut_name")

pred_joined$mut_name=NULL
corr_matrix <- cor(pred_joined,method = "spearman", use = "pairwise.complete.obs")
corr_matrix=corr_matrix[-5,-5]


pheatmap(corr_matrix,
         color = colorRampPalette(c("blue", "white", "red"))(50), 
         display_numbers = TRUE,      
         border_color = NA,          
         fontsize_row = 10,           
         fontsize_col = 10,           
         main = "Correlation Heatmap") 

melted_corr_matrix <- melt(corr_matrix)
melted_corr_matrix$value=abs(melted_corr_matrix$value)

ggplot(data = melted_corr_matrix, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Correlazione") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation Heatmap") +
  geom_text(aes(label = round(value, 2)), color = "black", size = 4)

my_order=c("mutpred", "provean", "sift", "polyphen", "cadd", "EVE", "alpha", "ESM1b",  "MITE")

melted_corr_matrix=melted_corr_matrix %>%
  mutate(
    Var1 = factor(Var1, levels = my_order),
    Var2 = factor(Var2, levels = my_order))

x=ggplot(data = melted_corr_matrix, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "#7fbf7b", high = "#af8dc3", mid = "white", 
                       midpoint = 0.57, limit = c(min(melted_corr_matrix$value),1), space = "Lab", 
                       name="Correlazione", na.value = "grey50" ) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation Heatmap") +
  geom_text(aes(label = round(value, 2)), color = "black", size = 4)

 pdf(paste0("Predictors_benchmark_RDS/Spearman_correlation_witheve","_",domain,".pdf"), height =4,width = 16)
x
dev.off()
```

Fig. 3B heatmap mean score predictor
```{r}
## DBD
fig="A5"
domain="DBD"

input=read.table(paste0("Data/",domain,"/Mutpred_prediction",domain,".txt"), row.names = 1,header=T)

p=list()
for(i in 1:6){
mut_pred=input[,i,drop=F]
mut_pred$residues=substr(rownames(mut_pred), 1, 4)
mut_pred$aa=substr(rownames(mut_pred), 5, 5)
mut_pred$order=substr(mut_pred$residues, 2,4)
mut_pred=mut_pred[order(mut_pred$order, decreasing=T),]
colnames(mut_pred)[1]="score"
mut_pred$residues=factor(mut_pred$residues, levels = rev(unique(mut_pred$residues)))
mut_pred$aa=factor(mut_pred$aa, levels = rev(sort))

library(tidyr)
z <- as.data.frame(complete(mut_pred, residues, aa))

plot=data.frame(residues=unique(z$residues), mean_score=sapply(unique(z$residues), function(x) mean(z[z$residues==x,"score"], na.rm=T)))

if(i==1){
limit=0.8
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)
} else if(i==2) {
limit=(-2.2)
colors=colorRampPalette(c("red2", 'red2', 'white', "darkblue", "darkblue"))(100)
} else if(i==3) {
limit=0.05
colors=colorRampPalette(c("red2", 'red2', 'white', "darkblue", "darkblue"))(100)
} else {
limit=0.5
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)
}

library(scales)
p[[i]]=ggplot(plot, aes(residues, "Cond")) + theme_void()+
  geom_tile(aes(fill = mean_score),size=1, color = "white") +
  scale_fill_gradientn(colors=colors
                       , values=rescale(c(min(z$score[!is.na(z$score)]),limit,max(z$score[!is.na(z$score)]))), na.value = "grey50") +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size=15,family="Courier",angle = 90, vjust = 0.5, hjust=1),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(.5,"cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size=15,family="Courier"))
}

pdf(paste0("Figure2/",fig,"_",domain,".pdf"), height =4,width = 16)
for(i in 1:length(p)){
  print(p[[i]])
}
dev.off()

## SAM

fig="A5"
domain="SAM"

input=read.table(paste0("Data/",domain,"/Mutpred_prediction",domain,".txt"), row.names = 1,header=T)
p=list()
for(i in 1:6){
mut_pred=input[,i,drop=F]
mut_pred$residues=substr(rownames(mut_pred), 1, 4)
mut_pred$aa=substr(rownames(mut_pred), 5, 5)
mut_pred$order=substr(mut_pred$residues, 2,4)
mut_pred=mut_pred[order(mut_pred$order, decreasing=T),]
colnames(mut_pred)[1]="score"
mut_pred$residues=factor(mut_pred$residues, levels = rev(unique(mut_pred$residues)))
mut_pred$aa=factor(mut_pred$aa, levels = rev(sort))


z <- as.data.frame(complete(mut_pred, residues, aa))
plot=data.frame(residues=unique(z$residues), mean_score=sapply(unique(z$residues), function(x) mean(z[z$residues==x,"score"], na.rm=T)))

if(i==1){
limit=0.8
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)
} else if(i==2) {
limit=(-2.2)
colors=colorRampPalette(c("red2", 'red2', 'white', "darkblue", "darkblue"))(100)
} else if(i==3) {
limit=0.05
colors=colorRampPalette(c("red2", 'red2', 'white', "darkblue", "darkblue"))(100)
} else {
limit=0.5
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)
}


p[[i]]=ggplot(plot, aes(residues, "Cond")) + theme_void()+
  geom_tile(aes(fill = mean_score),size=1, color = "white") +
  scale_fill_gradientn(colors=colors
                       , values=rescale(c(min(z$score[!is.na(z$score)]),limit,max(z$score[!is.na(z$score)]))), na.value = "grey50") +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size=15,family="Courier",angle = 90, vjust = 0.5, hjust=1),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(.5,"cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size=15,family="Courier"))
}

pdf(paste0("Figure2/",fig,"_",domain,".pdf"), height =4,width = 16)
for(i in 1:length(p)){
  print(p[[i]])
}
dev.off()

domain="DBD" #or SAM

input_pred=read.table(paste0(domain,"_prediction_complete_updated.txt"), row.names = 1,header=T)
cadd=input_pred[,c(7,8,9)]
rownames(cadd)=rownames(input_pred)
mut_pred=cadd
mut_pred$residues=substr(rownames(mut_pred), 1, 4)
mut_pred$aa=substr(rownames(mut_pred), 5, 5)
mut_pred$order=substr(mut_pred$residues, 2,4)
mut_pred=mut_pred[order(mut_pred$order, decreasing=T),]
colnames(mut_pred)[3]="score"
mut_pred$residues=factor(mut_pred$residues, levels = rev(unique(mut_pred$residues)))
mut_pred$aa=factor(mut_pred$aa, levels = rev(sort))


z <- as.data.frame(complete(mut_pred, residues, aa))
plot=data.frame(residues=unique(z$residues), mean_score=sapply(unique(z$residues), function(x) mean(z[z$residues==x,"score"], na.rm=T)))

limit=20
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)


p <- ggplot(plot, aes(residues, "CADD")) + theme_void() +
  geom_tile(aes(fill = mean_score), size = 1, color = "white") +
  scale_fill_gradientn(
    colors = colors,
    values = rescale(c(0, limit, max(z$score, na.rm = TRUE))),
    limits = c(0, max(z$score, na.rm = TRUE)),
    na.value = "grey50"
  ) +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 15, family = "Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(0.5, "cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size = 15, family = "Courier")
  )

pdf(paste0("figures/3B","_",domain,".pdf"), height =4,width = 16)
p
dev.off()

#value average position EVE

domain="SAM" #or DBD

input_pred=read.table(paste0(domain,"_prediction_complete_updated.txt"), row.names = 1,header=T)
pos=input_pred[,c(7,8)]
pos$mut_name=pos$mut
input_tot=read.table(paste0("input_con_eve_",domain,".tsv"), header=T)

input_joined=pos %>%
  left_join(input_tot %>% dplyr::select(mut_name, EVE), by = "mut_name")

rownames(input_joined)=input_joined$mut_name
input_joined$mut_name=NULL
input_joined$mut=NULL

pos=input_joined[,c(1,2)]

mut_pred=pos
mut_pred$residues=substr(rownames(mut_pred), 1, 4)
mut_pred$aa=substr(rownames(mut_pred), 5, 5)
mut_pred$order=substr(mut_pred$residues, 2,4)
mut_pred=mut_pred[order(mut_pred$order, decreasing=T),]
colnames(mut_pred)[2]="score"
mut_pred$residues=factor(mut_pred$residues, levels = rev(unique(mut_pred$residues)))
mut_pred$aa=factor(mut_pred$aa, levels = rev(sort))

z <- as.data.frame(complete(mut_pred, residues, aa))

plot=data.frame(residues=unique(z$residues), mean_score=sapply(unique(z$residues), function(x) mean(z[z$residues==x,"score"], na.rm=T)))

limit=0.44
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)

p <- ggplot(plot, aes(residues, "EVE")) + theme_void() +
  geom_tile(aes(fill = mean_score), size = 1, color = "white") +
  scale_fill_gradientn(
    colors = colors,
    values = rescale(c(0, limit, max(z$score, na.rm = TRUE))),
    limits = c(0, max(z$score, na.rm = TRUE)),
    na.value = "grey50"
  ) +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 15, family = "Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(0.5, "cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size = 15, family = "Courier")
  )


pdf(paste0("figures/3B_EVE","_",domain,".pdf"), height =4,width = 16)
p
dev.off()
```

Fig. 3A cumulative score
```{r}
sam=read.table("Data/SAM_mut_list_muscan_single_rep_sd_se.txt", header = T)
dbd=read.table("Data/DBD_mut_list_muscan_single_rep_sd_se.txt", header = T)


sub=dbd
sub=sub[rownames(sub)!="WT",]
sub$residues=substr(rownames(sub), 1, 4)
sub$aa=substr(rownames(sub), 5, 5)
sub$order=substr(sub$residues, 2,4)
sub=sub[order(sub$order, decreasing=T),]
sub$residues=factor(sub$residues, levels = rev(unique(sub$residues)))
sub_dbd=sub[,c("aa","logFC")]


sub=sam
sub=sub[rownames(sub)!="WT",]
sub$residues=substr(rownames(sub), 1, 4)
sub$aa=substr(rownames(sub), 5, 5)
sub$order=substr(sub$residues, 2,4)
sub=sub[order(sub$order, decreasing=T),]

sub$residues=factor(sub$residues, levels = rev(unique(sub$residues)))
sub_sam=sub[,c("aa","logFC")]

final=rbind(sub_dbd,sub_sam)
sort=sort(sapply(unique(final$aa), function(x) sum(final[final$aa==x, "logFC"],na.rm=T)))

sort=data.frame(aa=names(sort),values=sort)
sort$aa=factor(sort$aa,levels=(sort$aa))
write.table(sort, "Data/order_aa_heatmaps.txt",row.names = F,sep="\t",quote=F)


pdf(paste0("figures/","cumulative_score",".pdf"), height =6,width = 4)
ggplot(sort, aes(values, aa,group=1)) + theme_bw()+
  geom_point(size=3, color = "white") +
  geom_line()+
  ylab("") +
  xlab("") +
  theme(panel.grid = element_blank(),
    axis.text.x = element_text(size=15,family="Courier"),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.position = "none") +
  scale_x_continuous(position = "top")
dev.off()

```

Fig. 3C and Fig. 3D Polarity and Conservation
```{r}
fig="A2"
domain="DBD" #or SAM

clust=read.table(paste0("Data/Consurf_polarity_",domain,".txt"),header=T)

pdf(paste0("Figure2/",fig,"_",domain,".pdf"), height =4,width = 16)
ggplot(clust, aes(pos, "Cond")) + theme_void()+
  geom_tile(aes(fill = as.factor(polarity)),size=1, color = "white") +
  scale_fill_manual(values = c("darkorange4", "green4")) +
  ylab("") +
  xlab("") +
  theme(
    axis.text = element_blank(),
    legend.position = "none")
ggplot(clust, aes(pos, "Cond")) + theme_void()+
  geom_tile(aes(fill = as.factor(consurf)),size=1, color = "white") +
  scale_fill_manual(values =unlist(strsplit("#10C8D1,#8CFFFF,#D7FFFF,#EAFFFF,#FFFFFF,#FCEDF4,#FAC9DE,#F07DAB,#A02560",","))) +
  ylab("") +
  xlab("") +
  theme(
    axis.text = element_blank())
dev.off()
```

Fig. 3A Mite enrichment score
```{r}
domain="DBD" #or SAM
edger_scores=read.table(paste0("Data/",domain,"_mut_list_muscan_single_rep.txt"),sep="\t")

#preparing data for heatmap
sub=edger_scores
sub=sub[rownames(sub)!="WT",]
sub$residues=substr(rownames(sub), 1, 4)
sub$aa=substr(rownames(sub), 5, 5)
sub$order=substr(sub$residues, 2,4)
sub=sub[order(sub$order, decreasing=T),]

sub$residues=factor(sub$residues, levels = rev(unique(sub$residues)))
sort=read.table("Results/order_aa_heatmaps.txt",header=T)
sort=as.character(sort$aa)
sub$aa=factor(sub$aa, levels = rev(sort))


#we plot in blue the variants enriched in the converted and in red the variant enriched in not converted
pdf(paste0("figures/","Fig2A","_",domain,".pdf"), height =4,width = 16)
ggplot(sub, aes(aa, residues)) + theme_void()+
  geom_tile(aes(fill = logFC),size=1, color = "white") +
  scale_fill_gradientn(colors=colorRampPalette(c("darkblue", 'darkblue', 'white', "red", "red2"))(100),limits = c(-6, 6.80), na.value = "grey50") +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size=15,family="Courier",angle = 90, vjust = 0.5, hjust=1),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(.5,"cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size=15,family="Courier")) +
  #scale_x_discrete(position = "top")+
  coord_flip()
dev.off()
```

Fig. 4A score ranking
```{r}
domain="SAM" #DBD
clust=read.table(paste0("Data/",domain,"_mut_list_muscan_single_rep.txt"),header = T)

sub=clust[,"logFC",drop=F]
sub=sub[rownames(sub)!="WT",,drop=F]
sub$residues=substr(rownames(sub), 1, 4)
sub$aa_change=substr(rownames(sub), 5, 5)
sub$order=as.numeric(substr(sub$residues, 2,4))+43
sub$aa_base=substr(sub$residues, 1,1)
rownames(sub)=paste0(sub$aa_base,sub$order,sub$aa_change)
clust_sam=sub

domain="DBD" #SAM
clust=read.table(paste0("Data/",domain,"_mut_list_muscan_single_rep.txt"),header = T)

sub=clust[,"logFC",drop=F]
sub=sub[rownames(sub)!="WT",,drop=F]
sub$residues=substr(rownames(sub), 1, 4)
sub$aa_change=substr(rownames(sub), 5, 5)
sub$order=as.numeric(substr(sub$residues, 2,4))+43
sub$aa_base=substr(sub$residues, 1,1)
rownames(sub)=paste0(sub$aa_base,sub$order,sub$aa_change)

final=rbind(clust_sam[,"logFC",drop=F],sub[,"logFC",drop=F])
final$mut=rownames(final)
final$color=ifelse(final$mut%in%c("L514F","C522G","R304Q","R279Q","I550D", "L548N"),"Known Causative", "A")
final$color=ifelse(final$mut%in%c("R279S","R304T","L514D","C522D"),"New in known", final$color)
final$color=ifelse(final$mut%in%c("L531E","L531R","G297L","R311I"),"New hotsposts", final$color)
final$color=ifelse(final$mut%in%c("G293H","G310Q","T527I","H556A"),"Neutral", final$color)
final$color=ifelse(final$mut%in%c("WT","WildType"),"WT", final$color)
final$label=ifelse(final$color!="A", as.character(final$mut), "")
final$size=ifelse(final$color!="A", "YES","NO")
final=final[order(final$logFC),]
final$mut=factor(final$mut, levels=final$mut)
final$logFC_z <- (final$logFC - mean(final$logFC, na.rm = TRUE)) / sd(final$logFC, na.rm = TRUE)


pdf(paste0("figures/Score_ranking.pdf"), height =8,width =4 , useDingbats = F)
ggplot(final, aes(mut, logFC_z, fill=color,size=size,label=label)) + 
  geom_hline(yintercept = 0, linetype="dotted", color="grey20")+
  geom_point(pch=21, color="black") + theme_bw()+theme(axis.title = element_blank(), axis.ticks.x=element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.y = element_text(size=15, angle=90, hjust=0.5,family="Courier"), axis.text.x = element_blank(),legend.position = "none") +  scale_fill_manual(values=c("black","red2","darkorchid","orange2","blue2","green2")) + coord_cartesian(ylim=c(-4.25,4.25)) + scale_size_manual(values = c(0.5,3)) + geom_text(hjust=-0.5, vjust=0.5, size=3)
dev.off()
```

Fig. 4B IF, FACS and RNAseq quantification
```{r}
fig="3A"
levels=unlist(strsplit("WT - G293H - G310Q - R304Q - R279Q - R304T - R279S - R311I - G297L - T527I - H556A - L514F - C522G - L514D - C522D - L531E - L531R - L548N - I550D", " - "))

#IF
data=read.table("Paper_FP/Data/IF_quant.txt")
data$exp="IF"

#FACS
data_facs=readRDS("Data/mutants_FACS_data.rds")
data_facs_wt=readRDS("Data/wt_FACS_data.rds")
data_facs=rbind(data_facs, data_facs_wt)

threshold=quantile(data_facs[data_facs$sample=="WT","PE.H"],.25)
data_fin=data.frame(V1=unique(data_facs$sample), V2=sapply(unique(data_facs$sample), function(x) sum(data_facs[data_facs$sample==x, "PE.H"]>=threshold)/nrow(data_facs[data_facs$sample==x, ])*100), exp="FACS")
data_fin$V1=gsub("r304t","R304T",data_fin$V1)
data_fin$V2=ifelse(data_fin$V2>100,100,data_fin$V2)

#RNA
d=readRDS("Data/RNAseq_correlation.rds")
plot=rbind(data, data_fin,d)
plot$V1=factor(plot$V1,levels=levels)

#NATIVE

nat=read.table("Data/Native_quant.txt")
nat$V1=factor(nat$V1,levels=levels)
nat$V2=ifelse(nat$V2>4, 4, nat$V2)

pdf(paste0("Figure3/",fig,".pdf"), height =4,width = 16)
ggplot(plot[plot$exp!="RNA",], aes(V1, exp)) + theme_classic()+ theme(axis.title = element_blank(), axis.ticks = element_blank(), axis.text.x = element_text(angle=90,hjust=1, vjust=.5))+
  geom_tile(aes(fill = V2),size=1, color = "white") +
  scale_fill_distiller(palette = "OrRd", direction = 1)
ggplot(plot[plot$exp=="RNA",], aes(V1, exp)) + theme_classic()+ theme(axis.title = element_blank(), axis.ticks = element_blank(), axis.text.x = element_text(angle=90,hjust=1, vjust=.5))+
  geom_tile(aes(fill = V2),size=1, color = "white") +
  scale_fill_distiller(palette = "OrRd", direction = 1)
ggplot(nat, aes(V1, "Native")) + theme_classic()+ theme(axis.title = element_blank(), axis.ticks = element_blank(), axis.text.x = element_text(angle=90,hjust=1, vjust=.5))+
  geom_tile(aes(fill = V2),size=1, color = "white") +
  scale_fill_distiller(palette = "OrRd", direction = 1)
dev.off()
```

Fig. 4E heatmap
```{r}
fig="3D"
levels=unlist(strsplit("WT - G293H - G310Q - R304Q - R279Q - R304T - R279S - R311I - G297L - T527I - H556A - L514F - C522G - L514D - C522D - L531E - L531R - L548N - I550D - GFP", " - "))

mtx=read.table("Data/P63_Mutants_validation_final_merged_rawCountsWithAnnotations.txt",header=T,sep="\t",quote="",stringsAsFactors = F)

pjsamples=data.frame(unique_samples=colnames(mtx)[-c(1:2)], condition=ifelse(stringr::str_count(colnames(mtx)[-c(1:2)],"_")>1,sapply(strsplit(colnames(mtx)[-c(1,2)],"_"),"[[", 2),sapply(strsplit(colnames(mtx)[-c(1,2)],"_"),"[[", 1)))

mtx_red=mtx[,!grepl("NI|WTfir",colnames(mtx))]
pjsamples_red=pjsamples[pjsamples$unique_samples%in%colnames(mtx_red),]
pjsamples_red$condition=ifelse(grepl("WT",pjsamples_red$condition),"WT",pjsamples_red$condition)
pjsamples_red$condition=ifelse(grepl("GFP",pjsamples_red$condition),"GFP",pjsamples_red$condition)
cpm=cbind(mtx_red[,1:2], log2(cpm(mtx_red[,-c(1:2)])+1))
cpm$external_gene_name=ifelse(is.na(cpm$external_gene_name), cpm$ensembl_gene_id, cpm$external_gene_name)

avg=as.data.frame(sapply(unique(pjsamples_red$condition), function(x) rowMeans(cpm[colnames(cpm)%in%pjsamples_red[pjsamples_red$condition==x,1]])))
rownames(avg)=make.unique(cpm[,2])


# Subset for conversion signature
res=read.table("Results/WTvsGFP_results_annotated.txt",sep="\t",header=T)
res=res[res$diffEXP_WTvsGFP%in%c("UP","DOWN"),]
res$external_gene_name=ifelse(is.na(res$external_gene_name), res$ensembl_gene_id, res$external_gene_name)
sign=data.frame(row.names = res$external_gene_name, cond=res$diffEXP_WTvsGFP)
sign=sign[order(sign$cond,decreasing = T),,drop=F]
avg_red=avg[rownames(avg)%in%rownames(sign),]


#Create matrix
plot=as.matrix(avg_red[,match(levels,colnames(avg_red))])
plot=plot[match(rownames(sign),rownames(plot)),]

breaksList = c(seq(-2, -1, by = 0.01), seq(-0.99, -0.1, by = 0.01), seq(-0.09, 0.09, by=0.01), seq(0.1, 0.99, by=0.01), seq(1, 2, by = 0.01))

pdf(paste0("Figure3/",fig,"_UP_DOWN.pdf"), width=10,height=6)
pheatmap(plot, cluster_rows=F, show_rownames=F, border_color = NA, scale = "row", cluster_cols=F, color = colorRampPalette(c("#0571b0", "#92c5de", "#f7f7f7", "#f4a582", "#ca0020"))(length(breaksList)), breaks = breaksList,annotation_row = sign, angle_col = 0)
pheatmap(plot, cluster_rows=T, show_rownames=F, border_color = NA, scale = "row", cluster_cols=F, color = colorRampPalette(c("#0571b0", "#92c5de", "#f7f7f7", "#f4a582", "#ca0020"))(length(breaksList)), breaks = breaksList,annotation_row = sign, angle_col = 0)
pheatmap(plot[,colnames(plot)%in%c(levels[c(1:3,10,11,20)],"GFP")], cluster_rows=T, show_rownames=T, border_color = NA, scale = "row", cluster_cols=F, color = colorRampPalette(c("#0571b0", "#92c5de", "#f7f7f7", "#f4a582", "#ca0020"))(length(breaksList)), breaks = breaksList,annotation_row = sign, angle_col = 0,fontsize_row = 1)
dev.off()
```

Fig. 4D PCA 
```{r}
fig="S3A"
mtx=read.table("Data/P63_Mutants_validation_final_merged_rawCountsWithAnnotations.txt",header=T,sep="\t",quote="",stringsAsFactors = F)
mtx=mtx[,!grepl("WTfirst",colnames(mtx))]

pjsamples=data.frame(unique_samples=colnames(mtx)[-c(1:2)], condition=ifelse(stringr::str_count(colnames(mtx)[-c(1:2)],"_")>1,sapply(strsplit(colnames(mtx)[-c(1,2)],"_"),"[[", 2),sapply(strsplit(colnames(mtx)[-c(1,2)],"_"),"[[", 1)))

pjsamples$condition=ifelse(grepl("WT",pjsamples$condition),"WT",pjsamples$condition)
pjsamples$condition=ifelse(grepl("GFP",pjsamples$condition),"GFP",pjsamples$condition)

# PCA with Neutrals and no WT2
pjsamp_red=pjsamples[pjsamples$condition%in%c("WT","G293H","G310Q","T527I","H556A") & !grepl("WTsecond", pjsamples$unique_samples),]
cpm_red=cpm[,colnames(cpm)%in%pjsamp_red$unique_samples]

pca=prcomp(t(as.matrix(cpm_red)))
percentVar=pca$sdev^2/sum(pca$sdev^2)
d=merge(pca$x, pjsamp_red, by.x=0, by.y="unique_samples")
attr(d, "percentVar")=percentVar

ex=d
ex$class=ifelse(ex$condition%in%c("WT","G310Q"),ex$condition,"Neutral")

p1=ggplot(ex, aes(x=ex[,2], y=ex[,3], fill=class)) + geom_point(color="black", alpha=.7,size=5,pch=21) +
    theme_bw()+
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.text = element_text(size=10),
          axis.title = element_text(size=10),
          legend.title = element_text(size=10))+
    labs(fill="Condition")+
    xlab(paste0("PC1: ", round(attr(ex, "percentVar")[1] *100), "% variance"))+
    ylab(paste0("PC2: ", round(attr(ex, "percentVar")[2] *100), "% variance"))
p2=ggplot(ex, aes(x=ex[,2], y=ex[,3], fill=condition)) + geom_point(color="black", alpha=.7,size=5,pch=21) +
    theme_bw()+
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.text = element_text(size=10),
          axis.title = element_text(size=10),
          legend.title = element_text(size=10))+
    labs(fill="Condition")+
  scale_fill_manual(values = c('#e6194b', '#3cb44b', '#ffe119', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffd8b1', '#000075', '#808080', '#004600'))+
    xlab(paste0("PC1: ", round(attr(ex, "percentVar")[1] *100), "% variance"))+
    ylab(paste0("PC2: ", round(attr(ex, "percentVar")[2] *100), "% variance"))

pdf(paste0("SuppFigure3/",fig,"_Neut_noWT2.pdf"), height =6,width = 14)
grid.arrange(p1,p2,ncol=2,nrow=1, widths=c(1,1.1))
dev.off()
```

Fig. S4B PCA LoF-Neutral
```{r}
fig="S3A"

mtx=read.table("Data/P63_Mutants_validation_final_merged_rawCountsWithAnnotations.txt",header=T,sep="\t",quote="",stringsAsFactors = F)
mtx=mtx[,!grepl("WTfirst",colnames(mtx))]

pjsamples=data.frame(unique_samples=colnames(mtx)[-c(1:2)], condition=ifelse(stringr::str_count(colnames(mtx)[-c(1:2)],"_")>1,sapply(strsplit(colnames(mtx)[-c(1,2)],"_"),"[[", 2),sapply(strsplit(colnames(mtx)[-c(1,2)],"_"),"[[", 1)))

pjsamples$condition=ifelse(grepl("WT",pjsamples$condition),"WT",pjsamples$condition)
pjsamples$condition=ifelse(grepl("GFP",pjsamples$condition),"GFP",pjsamples$condition)
cpm=cbind(mtx[,1:2], log2(cpm(mtx[,-c(1:2)])+1))

pca=prcomp(t(as.matrix(cpm[,-c(1,2)])))
percentVar=pca$sdev^2/sum(pca$sdev^2)
d=merge(pca$x, pjsamples, by.x=0, by.y="unique_samples")
attr(d, "percentVar")=percentVar

ex=d
ex$class=ifelse(ex$condition%in%c("WT","G293H","G310Q","T527I","H556A"),"Neutral","LoF")

p1=ggplot(ex, aes(x=ex[,2], y=ex[,3], fill=class)) + geom_point(color="black", alpha=.7,size=5,pch=21) +
    theme_bw()+
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.text = element_text(size=10),
          axis.title = element_text(size=10),
          legend.title = element_text(size=10))+
    labs(fill="Condition")+
    xlab(paste0("PC1: ", round(attr(ex, "percentVar")[1] *100), "% variance"))+
    ylab(paste0("PC2: ", round(attr(ex, "percentVar")[2] *100), "% variance"))
p2=ggplot(ex, aes(x=ex[,2], y=ex[,3], fill=condition)) + geom_point(color="black", alpha=.7,size=5,pch=21) +
    theme_bw()+
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.text = element_text(size=10),
          axis.title = element_text(size=10),
          legend.title = element_text(size=10))+
    labs(fill="Condition")+
  scale_fill_manual(values = c('#e6194b', '#3cb44b', '#ffe119', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffd8b1', '#000075', '#808080', '#004600'))+
    xlab(paste0("PC1: ", round(attr(ex, "percentVar")[1] *100), "% variance"))+
    ylab(paste0("PC2: ", round(attr(ex, "percentVar")[2] *100), "% variance"))

pdf(paste0("SuppFigure3/",fig,".pdf"), height =6,width = 14)
grid.arrange(p1,p2,ncol=2,nrow=1, widths=c(1,1.1))
dev.off()

```

Fig. S4C GO ontology enrichment
```{r}

fig="S3A"
res=read.table("Results/WTvsGFP_results_annotated.txt",sep="\t",header=T)
sign=readRDS("Results/cluster_UP_DN_genes_GFPvsWT.rds")

plot=list()
list_genes=list(UP_c1=res[res$ensembl_gene_id%in%rownames(sign[sign$c=="c1",,drop=F]),2],DOWN_c2=res[res$ensembl_gene_id%in%rownames(sign[sign$c=="c2",,drop=F]),2], DOWN_c3=res[res$ensembl_gene_id%in%rownames(sign[sign$c=="c3",,drop=F]),2])

for(genes in 1:length(list_genes)){
dbs <- c("GO_Biological_Process_2023")
  enriched <- enrichr(list_genes[[genes]], dbs)
  
  p=list()
  for(i in 1:length(enriched)){
  enriched[[i]]$count=as.numeric(sapply(strsplit(as.character(enriched[[i]]$Overlap),"/"), "[[", 1))
  enriched[[i]]$total=as.numeric(sapply(strsplit(as.character(enriched[[i]]$Overlap),"/"), "[[", 2))
  enriched[[i]]$perc=enriched[[i]]$count/enriched[[i]]$total*100
  test=enriched[[i]]
  test=test[order(test$Adjusted.P.value, decreasing = F),]
  test$Term=factor(test$Term,levels=rev(test$Term))
  test$db=dbs[i]
  p[[i]]=test
  }
  kegg=p[[1]]
  kegg=kegg[order(kegg$Adjusted.P.value, decreasing = F),]
  kegg=kegg[1:5,c(1,3,12,4,13)]
  colnames(kegg)=c("Path","p.value","overlap","p.adj","db")
  
  data=kegg
  data$Path=factor(data$Path,levels=rev(data$Path))

plot[[genes]]=ggplot(data, aes(x = -log10(p.adj), y = Path, fill=overlap)) +
  theme_bw()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size=10),
        strip.background = element_blank(),
  strip.text = element_blank()) +
  ylab("") +
  xlab("")+
  labs(color = "")+
  geom_bar(color="black",stat = "identity",width=.7) +
  scale_fill_distiller(type = "seq", palette = "Reds",direction = 1)+
  geom_vline(xintercept = -log10(0.05),linetype="dashed",color="black")+
  ggtitle(names(list_genes)[genes]) +
  geom_text(aes(label=Path, x=0.2), hjust=0)
}

 pdf(paste0("SuppFigure3/",fig, "_GO_2.pdf"), height=4, width=8)
  for(p in 1:length(plot)){
 print(plot[[p]])
  }
  dev.off()
```

Fig. 5B featurePlot converted SCRAM-seq
```{r}
mite_red=readRDS("scRNA-seq_SMART/mite_seurat_sign_reduction.rds")

FeaturePlot(mite_red, features = c("TP63","KRT14","KRT5","ANPEP","SNAI2","ITGB4"))
ggsave("scRNA-seq_SMART/Feature_plots_genes_sign_red.pdf", height=10, width=10)
```

#Fig. 5Ea and 5Eb UMAP converted SCRAM-seq LoF-neutral
```{r}
dbd_mut=readRDS("scRNA-seq_SMART/DBD_final/DBD_BC_mut_assignment_with_effect.rds")
sam_mut=readRDS("scRNA-seq_SMART/SAM_final/SAM_BC_mut_assignment_with_effect.rds")

final_mut=rbind(dbd_mut, sam_mut)

b=merge(coord, final_mut, by.x=1, by.y="BC", all.x=T)
int_mut=unlist(strsplit("G293H - G310Q - R304Q - R279Q - R304T - R279S - R311I - G297L - T527I - H556A - L514F - C522G - L514D - C522D - L531E - L531R - L548N - I550D", " - "))
points=b[b$mut_name%in%int_mut,]

points_lof=b[b$Output=="LoF mutation",]
points_neu=b[b$Output=="Neutral mutation",]

g1=ggplot(coord, aes(x=UMAP_1, y=UMAP_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_lof, aes(fill=conversion), size=2, pch=21) + ggtitle("LoF mutations") +
  scale_fill_brewer(palette = "Dark2")

g2=ggplot(coord, aes(x=UMAP_1, y=UMAP_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_neu, aes(fill=conversion), size=2, pch=21) + ggtitle("Neutral mutations") +
  scale_fill_brewer(palette = "Dark2")

g1+g2
ggsave("5F_UMAP_all_mut_localized.pdf", height=8, width=16)

lof_plot=list()
colors=c('#1b9e77','#d95f02','#7570b3')
names(colors)=sort(unique(coord$conversion))

for(i in unique(coord$conversion)){
value=colors[i]
lof_plot[[i]]=ggplot(coord, aes(x=UMAP_1, y=UMAP_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_lof[points_lof$conversion==i,], aes(fill=conversion), size=2, pch=21) + ggtitle("LoF mutations") +
  scale_fill_manual(values=value)
}

lof_plot[[1]]+lof_plot[[2]]+lof_plot[[3]]
ggsave("scRNA-seq_SMART/5F_UMAP_all_mut_divided_LoF.pdf", height=6, width=16)


points_lof=b[b$Output=="LoF mutation",]
points_neu=b[b$Output=="Neutral mutation",]

g1=ggplot(coord, aes(x=UMAP_1, y=UMAP_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_lof, aes(fill=conversion), size=2, pch=21) + ggtitle("LoF mutations") +
  scale_fill_brewer(palette = "Dark2")

g2=ggplot(coord, aes(x=UMAP_1, y=UMAP_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_neu, aes(fill=conversion), size=2, pch=21) + ggtitle("Neutral mutations") +
  scale_fill_brewer(palette = "Dark2")

g1+g2
ggsave("5F_UMAP_all_mut_localized.pdf", height=8, width=16)

lof_plot=list()
colors=c('#1b9e77','#d95f02','#7570b3')
names(colors)=sort(unique(coord$conversion))

for(i in unique(coord$conversion)){
value=colors[i]
lof_plot[[i]]=ggplot(coord, aes(x=UMAP_1, y=UMAP_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_lof[points_lof$conversion==i,], aes(fill=conversion), size=2, pch=21) + ggtitle("LoF mutations") +
  scale_fill_manual(values=value)
}

lof_plot[[1]]+lof_plot[[2]]+lof_plot[[3]]
ggsave("5F_UMAP_all_mut_divided_LoF.pdf", height=6, width=16)

lof_plot=list()
colors=c('#1b9e77','#d95f02','#7570b3')
names(colors)=sort(unique(coord$conversion))

for(i in unique(coord$conversion)){
value=colors[i]
lof_plot[[i]]=ggplot(coord, aes(x=UMAP_1, y=UMAP_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_neu[points_neu$conversion==i,], aes(fill=conversion), size=2, pch=21) + ggtitle("Neutral mutations") +
  scale_fill_manual(values=value)
}

lof_plot[[1]]+lof_plot[[2]]+lof_plot[[3]]
ggsave("scRNA-seq_SMART/5F_UMAP_all_mut_divided_Neutral.pdf", height=6, width=16)

```

Fig. 5F converted SCRAM-seq mean score
```{r}
domain="SAM" #or DBD

sam_scores=readRDS("Results/SAM_final_score.rds")
sam_scores=sam_scores[rownames(sam_scores)!="WT",,drop=F]
rownames(sam_scores)=paste0(substr(rownames(sam_scores),1,1),as.numeric(substr(rownames(sam_scores),2,4))+43,substr(rownames(sam_scores),5,5))
sam_mut=readRDS("scRNA-seq_SMART/SAM_final/SAM_BC_mut_assignment_with_effect.rds")
fin_sam=merge(sam_mut[sam_mut$mut_name!="NA",], sam_scores, by.x="mut_name", by.y=0, all.x=T)

dbd_scores=readRDS("Results/DBD_final_score.rds")
dbd_scores=dbd_scores[rownames(dbd_scores)!="WT",,drop=F]
rownames(dbd_scores)=paste0(substr(rownames(dbd_scores),1,1),as.numeric(substr(rownames(dbd_scores),2,4))+43,substr(rownames(dbd_scores),5,5))
dbd_mut=readRDS("scRNA-seq_SMART/DBD_final/DBD_BC_mut_assignment_with_effect.rds")
fin_dbd=merge(dbd_mut[dbd_mut$mut_name!="NA",], dbd_scores, by.x="mut_name", by.y=0, all.x=T)

fin=rbind(fin_dbd, fin_sam)

df_dbd=fin %>%
  filter(str_starts(BC, "DBD_")) 
df_dbd= df_dbd %>%
  filter(!is.na(Output)) %>%
  group_by(mut_name,Output) %>%
  summarise(final_score=mean(final_score, na.rm=TRUE), .groups = "drop") 
df_dbd$residues=substr(df_dbd$mut_name, 1, 4)
df_dbd$aa=substr(df_dbd$mut_name, 5, 5)
df_dbd$order=substr(df_dbd$residues, 2,4)
df_dbd=df_dbd[order(df_dbd$order, decreasing=T),]
colnames(df_dbd)[1]="score"
df_dbd$residues=factor(df_dbd$residues, levels = rev(unique(df_dbd$residues)))
sort=sort=read.table("Paper_FP/Results//order_aa_heatmaps.txt", header = T)
sort=as.character(sort$aa)
df_dbd$aa=factor(df_dbd$aa, levels = rev(sort))

df_sam=fin %>%
  filter(str_starts(BC, "SAM_"))

df_sam= df_sam %>%
  filter(!is.na(Output)) %>%
  group_by(mut_name,Output) %>%
  summarise(final_score=mean(final_score, na.rm=TRUE), .groups = "drop") 


df_sam$residues=substr(df_sam$mut_name, 1, 4)
df_sam$aa=substr(df_sam$mut_name, 5, 5)
df_sam$order=substr(df_sam$residues, 2,4)
df_sam=df_sam[order(df_sam$order, decreasing=T),]
colnames(df_sam)[1]="score"
df_sam$residues=factor(df_sam$residues, levels = rev(unique(df_sam$residues)))
sort=sort=read.table("Results/order_aa_heatmaps.txt", header = T)
sort=as.character(sort$aa)
df_sam$aa=factor(df_sam$aa, levels = rev(sort))
sub=df_sam

sub=sub[order(sub$order, decreasing=T),]
sub$residues=as.character(sub$residues)
sort=sort=read.table("Results/order_aa_heatmaps.txt", header = T)
sort=as.character(sort$aa)
sub$aa=factor(sub$aa, levels = rev(sort))
sub$residues=factor(sub$residues, levels = unique(sub$residues))


plot <- data.frame(
  residues = names(tapply(sub$final_score, sub$residues, mean, na.rm = TRUE)),
  mean_score = as.numeric(tapply(sub$final_score, sub$residues, mean, na.rm = TRUE)))

plot$residues=factor(plot$residues, levels = rev(unique(plot$residues)))

pdf(paste0("figures/SCRAM_mean_score","_",domain,".pdf"), height =4,width = 16)
ggplot(plot, aes(residues, "Mean")) + theme_void()+
  geom_tile(aes(fill = mean_score),size=1, color = "white") +
  scale_fill_gradientn(colors=colorRampPalette(c("red2", 'red', 'white', "darkblue", "darkblue"))(100),limits = c(-1.75, 1.75), na.value = "grey50") +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size=15,family="Courier",angle = 90, vjust = 0.5, hjust=1),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(.5,"cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size=15,family="Courier"))
dev.off()
```

Fig. 5D density UMAP converted SCRAM-seq 
```{r}
mite_red=readRDS("scRNA-seq_SMART/mite_seurat_sign_reduction.rds")
coord=merge(mite_red@meta.data, mite_red@reductions$umap@cell.embeddings, by=0)

wt=read.table("Results/WTvsGFP_results_annotated.txt",sep="\t",header=T)
wt=wt[wt$diffEXP_WTvsGFP=="UP",]
wt=wt[order(wt$logFC_WTvsGFP,decreasing = T),]

genes=wt$external_gene_name[1:100]

mite_red <- AddModuleScore(
  object = mite_red,
  features = list(genes),
  ctrl = 5,
  name = 'TP63_sign'
)
df <- FetchData(mite_red, vars = c("umap_1", "umap_2", "TP63_sign1"))
df$group <- ifelse(df$TP63_sign1 >= 0, "High TP63Sign", "Low TP63Sign")


p=ggplot() +
  geom_point(data = df, aes(umap_1, umap_2), color = "grey90", size = 0.3) +
  stat_density_2d_filled(
    data = df %>% filter(group == "Low TP63Sign"),
    aes(x = umap_1, y = umap_2, fill = after_stat(level)),
    geom = "polygon", color = "white"
  ) +
 scale_fill_viridis_d(option = "inferno", name = "Density") +
  coord_cartesian(
    xlim = c(min(df$umap_1) - 1, max(df$umap_1) + 1),
    ylim = c(min(df$umap_2) - 1, max(df$umap_2) + 1)
  ) +
  theme_bw() 
p2=ggplot() +
  geom_point(data = df, aes(umap_1, umap_2), color = "grey90", size = 0.3) +
  stat_density_2d_filled(
    data = df %>% filter(group == "High TP63Sign"),
    aes(x = umap_1, y = umap_2, fill = after_stat(level)),
    geom = "polygon", color = "white"
  ) +
 scale_fill_viridis_d(option = "inferno", name = "Density") +
  coord_cartesian(
    xlim = c(min(df$umap_1) - 1, max(df$umap_1) + 1),
    ylim = c(min(df$umap_2) - 1, max(df$umap_2) + 1)
  ) +
  theme_bw() 
  
pdf("figures/density_umap_covertedscram_high.pdf", height = 8, width = 12)
p1+p2
dev.off()
```

Fig. S5B Correlation Spearman SCRAM-seq converted vs all predictors
```{r}
#scram vs all predictors
old_mite=readRDS("scRNA-seq_SMART/concordant_mut_SCRAM_MITE_metadata.rds")
old_mite=old_mite[old_mite$orig.ident=="DBD",] #SAM
b_clean <- old_mite[!is.na(old_mite$mut_name) & old_mite$mut_name != "", ]

df_summary <- b_clean %>%
  group_by(mut_name) %>%  
  summarise(
    mean_TP63_sign = mean(TP63_sign1, na.rm = TRUE),
    occurrence = n(),            
    .groups = "drop")

file_path <- "Predictors_benchmark_RDS/Output MITE and predictors.xlsx"  
dbd <- read_excel(file_path, sheet = 1)
dbd$mut_name=dbd$Variant  

dbd_allpre_mite<- dbd %>%
  inner_join(
    df_summary %>% dplyr::select(mut_name, mean_TP63_sign),
    by = "mut_name"
  )
dbd_allpre_mite$mutant_P53=NULL
dbd_allpre_mite$P53_DMS_score=NULL
dbd_allpre_mite$Variant=NULL
dbd_allpre_mite$Codon=NULL
dbd_allpre_mite$Output=NULL
dbd_allpre_mite$alpha=as.numeric(dbd_allpre_mite$alpha)
dbd_allpre_mite$papi=NULL
colnames(dbd_allpre_mite)=c("mutpred","provean","sift","polyphen","alpha","cadd","ESM1b","EVE","MITE","mut_name","SCRAM_c_DBD")

all_pre=dbd_allpre_mite[,!colnames(dbd_allpre_mite) %in% "mut_name" ]
all_pre <- as.data.frame(lapply(all_pre, as.numeric))

corr_matrix <- cor(all_pre, method = "spearman", use = "pairwise.complete.obs")
print(round(corr_matrix, 3))

melted_corr <- melt(corr_matrix)
melted_corr$value=abs(melted_corr$value)

ggplot(data = melted_corr, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "#7fbf7b", high = "#af8dc3", mid = "white", midpoint = 0.57, limit = c(min(melted_corr$value),1), space = "Lab", name="Correlazione", na.value = "grey50" ) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation Heatmap") +
  geom_text(aes(label = round(value, 2)), color = "black", size = 4)

ggsave("figures/Correlation_spearman_SCRAMc_DBD_MITE_allpredictors_avg_new.pdf", height=8, width=12)
ggsave("figures/Correlation_spearman_SCRAMc_SAM_MITE_allpredictors_avg_new.pdf", height=8, width=12)
```

Fig. S5C heatmap SCRAM converted SAM and DBD
```{r}
domain="SAM" #or DBD

sam_scores=readRDS("Paper_FP/Results/SAM_final_score.rds")
sam_scores=sam_scores[rownames(sam_scores)!="WT",,drop=F]
rownames(sam_scores)=paste0(substr(rownames(sam_scores),1,1),as.numeric(substr(rownames(sam_scores),2,4))+43,substr(rownames(sam_scores),5,5))
sam_mut=readRDS("scRNA-seq_SMART/SAM_final/SAM_BC_mut_assignment_with_effect.rds")
fin_sam=merge(sam_mut[sam_mut$mut_name!="NA",], sam_scores, by.x="mut_name", by.y=0, all.x=T)

dbd_scores=readRDS("Project_MITEseq/Paper/Results/DBD_final_score.rds")
dbd_scores=dbd_scores[rownames(dbd_scores)!="WT",,drop=F]
rownames(dbd_scores)=paste0(substr(rownames(dbd_scores),1,1),as.numeric(substr(rownames(dbd_scores),2,4))+43,substr(rownames(dbd_scores),5,5))
dbd_mut=readRDS("Project_MITEseq/scRNA-seq_SMART/DBD_final/DBD_BC_mut_assignment_with_effect.rds")
fin_dbd=merge(dbd_mut[dbd_mut$mut_name!="NA",], dbd_scores, by.x="mut_name", by.y=0, all.x=T)

fin=rbind(fin_dbd, fin_sam)

df_sam=fin %>%
  filter(str_starts(BC, "SAM_"))

df_sam= df_sam %>%
  filter(!is.na(Output)) %>%
  group_by(mut_name,Output) %>%
  summarise(final_score=mean(final_score, na.rm=TRUE), .groups = "drop") 

df_sam$residues=substr(df_sam$mut_name, 1, 4)
df_sam$aa=substr(df_sam$mut_name, 5, 5)
df_sam$order=substr(df_sam$residues, 2,4)
df_sam=df_sam[order(df_sam$order, decreasing=T),]
colnames(df_sam)[1]="score"
df_sam$residues=factor(df_sam$residues, levels = rev(unique(df_sam$residues)))
sort=sort=read.table("Results//order_aa_heatmaps.txt", header = T)
sort=as.character(sort$aa)
df_sam$aa=factor(df_sam$aa, levels = rev(sort))

z <- as.data.frame(complete(df_sam, residues, aa))

pdf(paste0("figures/Heatmap_scores_SCRAM_",domain,".pdf"), height =4,width = 16)
ggplot(z, aes(aa, residues)) + theme_void()+
  geom_tile(aes(fill = final_score),size=1, color = "white") +
  scale_fill_gradientn(colors=colorRampPalette(rev(c("darkblue", 'blue', 'white', "red", "red2")))(100)
                       , values=rescale(c(min(z$final_score[!is.na(z$final_score)]),0,max(z$final_score[!is.na(z$final_score)]))), na.value = "grey50") +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size=15,family="Courier",angle = 90, vjust = 0.5, hjust=1),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(.5,"cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size=15,family="Courier")) +
  coord_flip()
dev.off()
```


Fig. S5D and S5G confusions plot 
```{r}
file_path <- "Predictors_benchmark_RDS/Output MITE and predictors.xlsx"
sam <- read_excel(file_path, sheet = 2)
sam$mut_name=sam$Variant  

old_mite=readRDS("scRNA-seq_SMART/concordant_mut_SCRAM_MITE_metadata.rds")
old_mite=old_mite[old_mite$orig.ident=="SAM",]
b_clean <- old_mite[!is.na(old_mite$mut_name) & old_mite$mut_name != "", ]

scram_and_mite<- old_mite %>%
  inner_join(
    sam %>% dplyr::select(mut_name, MITE,mutpred,provean,sift,polyphen,alpha,cadd, ESM1b,EVE,Output),
    by = "mut_name"
  )
scram_and_mite$output_mite=scram_and_mite$Output.y
clean_df <- scram_and_mite %>%
  dplyr::select(mut_name, MITE, TP63_sign1,mutpred,provean,sift,polyphen,alpha,cadd,ESM1b,EVE,output_mite)

clean_df$output_old_scram=ifelse(clean_df$TP63_sign1>=0,"Neutral","LoF")
clean_df$output_mite=ifelse(clean_df$output_mite=="PAV","LoF","Neutral")

predictors=c("MITE","mutpred", "provean", "sift", "polyphen", "alpha", "cadd", "ESM1B", "eve")
output_dir="temp/"


df <- data.frame(
  Variant = clean_df$mut_name,
  MITE = clean_df$output_mite,
  SCRAM_c = clean_df$output_old_scram,
  mutpred = ifelse(clean_df$mutpred>=0.8,"LoF","Neutral"),
  provean = ifelse(clean_df$provean<=-2,"LoF","Neutral"),
  sift = ifelse(clean_df$sift<=0.05,"LoF","Neutral"),
  polyphen = ifelse(clean_df$polyphen>=0.5,"LoF","Neutral"),
  alpha = ifelse(clean_df$alpha>=0.56,"LoF","Neutral"),
  cadd = ifelse(clean_df$cadd>=20,"LoF","Neutral"),
  ESM1B = ifelse(clean_df$ESM1b<=-7.5,"LoF","Neutral"),
  eve = ifelse(clean_df$EVE>=0.44,"LoF","Neutral"))


df=df[!is.na(df$MITE),]
df=df %>% distinct(Variant, .keep_all = TRUE)

for (pred in predictors) {
  conf_tab <- df %>% 
    dplyr::count(SCRAM_c, !!sym(pred), name = "n") %>%
    rename(Predictor_class = !!sym(pred))

  p <- ggplot(conf_tab, aes(x = Predictor_class, y = SCRAM_c, fill = n)) +
    geom_tile(color = "white") +
    geom_text(aes(label = n), size = 5, color = "black", fontface = "bold") +
    scale_fill_gradient(
      low = "white",
      high = "#377eb8",
      limits = c(0, max(conf_tab$n)),
      expand = c(0, 0)
    ) +
    theme_minimal(base_size = 14) +
    labs(
      title = paste("Concordanza SCRAM (converted) vs", pred),
      x = paste("Classificazione", pred),
      y = "Classificazione SCRAM (converted)",
      fill = "n varianti"
    ) +
    theme(
      panel.grid = element_blank(),
      axis.text = element_text(size = 12, color = "black"),
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
    )

  filename <- paste0(output_dir, "confusionplot_conversionSCRAMvs_", pred, ".pdf")
  ggsave(filename, plot = p, height = 5, width = 8)

  cat("✅ Salvato:", filename, "\n")
}

file_path <- "Predictors_benchmark_RDS/Output MITE and predictors.xlsx"  
dbd <- read_excel(file_path, sheet = 1)
dbd$mut_name=dbd$Variant  

old_mite=readRDS("scRNA-seq_SMART/concordant_mut_SCRAM_MITE_metadata.rds")
old_mite=old_mite[old_mite$orig.ident=="DBD",]
b_clean <- old_mite[!is.na(old_mite$mut_name) & old_mite$mut_name != "", ]

scram_and_mite<- b_clean %>%
  inner_join(
    dbd %>% dplyr::select(mut_name, MITE,Output),
    by = "mut_name"
  )

scram_and_mite$output_mite=scram_and_mite$Output.y

clean_df <- scram_and_mite %>% dplyr::select(mut_name, MITE, TP63_sign1,output_mite)
clean_df$output_old_scram=ifelse(clean_df$TP63_sign1>=0,"Neutral","LoF")
clean_df$output_mite=ifelse(clean_df$output_mite=="PAV","LoF","Neutral")

df <- data.frame(
  Variant = clean_df$mut_name,
  MITE = clean_df$output_mite,
  SCRAM_c = clean_df$output_old_scram)

df=df[!is.na(df$MITE),]
df=df %>% distinct(Variant, .keep_all = TRUE)
  conf_tab <- df %>% 
    dplyr::count(SCRAM_c,MITE, name = "n")

  p <- ggplot(conf_tab, aes(x = MITE, y = SCRAM_c, fill = n)) +
    geom_tile(color = "white") +
    geom_text(aes(label = n), size = 5, color = "black", fontface = "bold") +
    scale_fill_gradient(
      low = "white",
      high = "#377eb8",
      limits = c(0, max(conf_tab$n)),
      expand = c(0, 0)
    ) +
    theme_minimal(base_size = 14) +
    labs(
      title = paste("Concordanza DBD SCRAM ( converted) vs", "MITE"),
      x = paste("Classificazione", "MITE"),
      y = "Classificazione SCRAM ( converted)",
      fill = "n varianti"
    ) +
    theme(
      panel.grid = element_blank(),
      axis.text = element_text(size = 12, color = "black"),
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
    )

  filename <- paste0(output_dir, "confusionplot_DBDconversionSCRAMvs_", "MITE", ".pdf")
  ggsave(filename, plot = p, height = 5, width = 8)
  
file_path <- "Predictors_benchmark_RDS/Output MITE and predictors.xlsx"  
dbd <- read_excel(file_path, sheet = 1)
dbd$mut_name=dbd$Variant  

old_mite=readRDS("scRNA-seq_SMART/concordant_mut_SCRAM_MITE_metadata.rds")
old_mite=old_mite[old_mite$orig.ident=="DBD",]
b_clean <- old_mite[!is.na(old_mite$mut_name) & old_mite$mut_name != "", ]

scram_and_mite<- b_clean %>%
  inner_join(
    dbd %>% dplyr::select(mut_name, MITE,mutpred,provean,sift,polyphen,papi,alpha,cadd, ESM1b,EVE,Output),
    by = "mut_name"
  )
scram_and_mite$output_mite=scram_and_mite$Output.y

clean_df <- scram_and_mite %>%
  dplyr::select(mut_name, MITE, TP63_sign1,mutpred,provean,sift,polyphen,papi,alpha,cadd,ESM1b,EVE,output_mite)

clean_df$output_old_scram=ifelse(clean_df$TP63_sign1>=0,"Neutral","LoF")
clean_df$output_mite=ifelse(clean_df$output_mite=="PAV","LoF","Neutral")

predictors=c("mutpred", "provean", "sift", "polyphen", "alpha", "cadd", "ESM1B", "eve")
output_dir="figures/"

df <- data.frame(
  Variant = clean_df$mut_name,
  MITE = clean_df$output_mite,
  SCRAM_c = clean_df$output_old_scram,
  mutpred = ifelse(clean_df$mutpred>=0.8,"LoF","Neutral"),
  provean = ifelse(clean_df$provean<=-2,"LoF","Neutral"),
  sift = ifelse(clean_df$sift<=0.05,"LoF","Neutral"),
  polyphen = ifelse(clean_df$polyphen>=0.5,"LoF","Neutral"),
  papi = clean_df$papi,
  alpha = ifelse(clean_df$alpha>=0.56,"LoF","Neutral"),
  cadd = ifelse(clean_df$cadd>=20,"LoF","Neutral"),
  ESM1B = ifelse(clean_df$ESM1b<=-7.5,"LoF","Neutral"),
  eve = ifelse(clean_df$EVE>=0.44,"LoF","Neutral"))


df=df[!is.na(df$MITE),]
df$papi=NULL
df=df %>% distinct(Variant, .keep_all = TRUE)

for (pred in predictors) {
  conf_tab <- df %>% 
    dplyr::count(SCRAM_c, !!sym(pred), name = "n") %>%
    rename(Predictor_class = !!sym(pred))

  p <- ggplot(conf_tab, aes(x = Predictor_class, y = SCRAM_c, fill = n)) +
    geom_tile(color = "white") +
    geom_text(aes(label = n), size = 5, color = "black", fontface = "bold") +
    scale_fill_gradient(
      low = "white",
      high = "#377eb8",
      limits = c(0, max(conf_tab$n)),
      expand = c(0, 0)
    ) +
    theme_minimal(base_size = 14) +
    labs(
      title = paste("Concordanza SCRAM (converted) vs", pred),
      x = paste("Classificazione", pred),
      y = "Classificazione SCRAM (converted)",
      fill = "n varianti"
    ) +
    theme(
      panel.grid = element_blank(),
      axis.text = element_text(size = 12, color = "black"),
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
    )

  filename <- paste0(output_dir, "confusionplot_DBDconversionSCRAMvs_", pred, ".pdf")
  ggsave(filename, plot = p, height = 5, width = 8)

  cat("✅ Salvato:", filename, "\n")
}
```


Fig.S5C heatmap SCRAM convertito SAM and DBD
```{r}
domain="DBD" #SAM

sam_scores=readRDS("Results/SAM_final_score.rds")
sam_scores=sam_scores[rownames(sam_scores)!="WT",,drop=F]
rownames(sam_scores)=paste0(substr(rownames(sam_scores),1,1),as.numeric(substr(rownames(sam_scores),2,4))+43,substr(rownames(sam_scores),5,5))
sam_mut=readRDS("scRNA-seq_SMART/SAM_final/SAM_BC_mut_assignment_with_effect.rds")
fin_sam=merge(sam_mut[sam_mut$mut_name!="NA",], sam_scores, by.x="mut_name", by.y=0, all.x=T)

dbd_scores=readRDS("Results/DBD_final_score.rds")
dbd_scores=dbd_scores[rownames(dbd_scores)!="WT",,drop=F]
rownames(dbd_scores)=paste0(substr(rownames(dbd_scores),1,1),as.numeric(substr(rownames(dbd_scores),2,4))+43,substr(rownames(dbd_scores),5,5))
dbd_mut=readRDS("scRNA-seq_SMART/DBD_final/DBD_BC_mut_assignment_with_effect.rds")
fin_dbd=merge(dbd_mut[dbd_mut$mut_name!="NA",], dbd_scores, by.x="mut_name", by.y=0, all.x=T)

fin=rbind(fin_dbd, fin_sam)
df_dbd=fin %>%
  filter(str_starts(BC, "DBD_"))

df_dbd= df_dbd %>%
  filter(!is.na(Output)) %>%
  group_by(mut_name,Output) %>%
  summarise(final_score=mean(final_score, na.rm=TRUE), .groups = "drop") 

df_dbd$residues=substr(df_dbd$mut_name, 1, 4)
df_dbd$aa=substr(df_dbd$mut_name, 5, 5)
df_dbd$order=substr(df_dbd$residues, 2,4)
df_dbd=df_dbd[order(df_dbd$order, decreasing=T),]
colnames(df_dbd)[1]="score"
df_dbd$residues=factor(df_dbd$residues, levels = rev(unique(df_dbd$residues)))
sort=sort=read.table("/Results/order_aa_heatmaps.txt", header = T)
sort=as.character(sort$aa)
df_dbd$aa=factor(df_dbd$aa, levels = rev(sort))

z <- as.data.frame(complete(df_dbd, residues, aa))

pdf(paste0("figures/Heatmap_scores_SCRAM_",domain,".pdf"), height =4,width = 16)
ggplot(z, aes(aa, residues)) + theme_void()+
  geom_tile(aes(fill = final_score),size=1, color = "white") +
  scale_fill_gradientn(colors=colorRampPalette(rev(c("darkblue", 'blue', 'white', "red", "red2")))(100)
                       , values=rescale(c(min(z$final_score[!is.na(z$final_score)]),0,max(z$final_score[!is.na(z$final_score)]))), na.value = "grey50") +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size=15,family="Courier",angle = 90, vjust = 0.5, hjust=1),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(.5,"cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size=15,family="Courier")) +
  coord_flip()
dev.off()
```

Fig. S5F ROC CURVE MITE SCRAM-seq converted and SCRAM-seq converted vs all predictors
```{r}
old_mite=readRDS("scRNA-seq_SMART/concordant_mut_SCRAM_MITE_metadata.rds")
old_mite=old_mite[old_mite$orig.ident=="SAM",]
b_clean <- old_mite[!is.na(old_mite$mut_name) & old_mite$mut_name != "", ]

scram_allpre_mite=rbind(dbd_allpre_mite,sam_allpre_mite)
scram_allpre_mite$output_old_scram=ifelse(scram_allpre_mite$mean_TP63_sign>=0,"Neutral","LoF")

df_summary=scram_allpre_mite
df_summary$output_mite=df_summary$Output
df_summary$output_old_scram=ifelse(df_summary$mean_TP63_sign>=0,"Neutral","LoF")

df_summary$output_mite=ifelse(df_summary$output_mite=="PAV","LoF","Neutral")

scram_c_mite <- df_summary %>% dplyr::select(mut_name,mean_TP63_sign, output_old_scram,output_mite)

df_pred <- scram_c_mite %>%
  mutate(
    mite_label = ifelse(output_mite == "LoF", 1, 0),
  )

roc_list <- list(
 "MITE"  <- roc(df_pred$mite_label, df_pred$mite_label,      direction = "<"),
"SCRAM_c" <- roc(df_pred$mite_label, -df_pred$mean_TP63_sign,  direction = "<"))

names(roc_list) <- c("MITE", "SCRAM_c")
auc_values <- sapply(roc_list, auc)
names(roc_list) <- paste0(names(roc_list), " (AUC=", sprintf("%.2f", auc_values), ")")

# Plot con AUC in legenda
ggroc(roc_list, legacy.axes = TRUE) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
  theme_minimal(base_size = 14) +
  labs(
    title = "ROC — Confronto SCRAM conversion SAM/DBD vs MITE (riferimento)",
    x = "False Positive Rate (1 - Specificità)",
    y = "True Positive Rate (Sensibilità)",
    color = "Predictor (AUC)"
  ) +
  theme(legend.position = "right")

ggsave("figures/ROC_MITE_cSCRAM_SAM&DBD_avg.pdf", height=5, width=8)

old_mite=readRDS("scRNA-seq_SMART/concordant_mut_SCRAM_MITE_metadata.rds")
old_mite=old_mite[old_mite$orig.ident=="DBD",]
b_clean <- old_mite[!is.na(old_mite$mut_name) & old_mite$mut_name != "", ]

df_summary <- b_clean %>%
  group_by(mut_name) %>% 
  summarise(
    mean_TP63_sign = mean(TP63_sign1, na.rm = TRUE),
    occurrence = n(),            
    .groups = "drop"
  )

file_path <- "Predictors_benchmark_RDS/Output MITE and predictors.xlsx"  
dbd <- read_excel(file_path, sheet = 1)
dbd$mut_name=dbd$Variant  

dbd_allpre_mite<- dbd %>%
  inner_join(
    df_summary %>% dplyr::select(mut_name, mean_TP63_sign),
    by = "mut_name"
  )
dbd_allpre_mite$mutant_P53=NULL
dbd_allpre_mite$P53_DMS_score=NULL

#SAM SCRAM_c

old_mite=readRDS("scRNA-seq_SMART/concordant_mut_SCRAM_MITE_metadata.rds")
old_mite=old_mite[old_mite$orig.ident=="SAM",]
b_clean <- old_mite[!is.na(old_mite$mut_name) & old_mite$mut_name != "", ]

df_summary <- b_clean %>%
  group_by(mut_name) %>%  #
  summarise(
    mean_TP63_sign = mean(TP63_sign1, na.rm = TRUE),
    occurrence = n(),             
    .groups = "drop"
  )

file_path <- "Predictors_benchmark_RDS/Output MITE and predictors.xlsx"  
sam <- read_excel(file_path, sheet = 2)
sam$mut_name=sam$Variant  

sam_allpre_mite<- sam %>%
  inner_join(
    df_summary %>% dplyr::select(mut_name, mean_TP63_sign),
    by = "mut_name"
  )

scram_allpre_mite=rbind(dbd_allpre_mite,sam_allpre_mite)

scram_allpre_mite$output_old_scram=ifelse(scram_allpre_mite$mean_TP63_sign>=0,"Neutral","LoF")

df_pred <- scram_allpre_mite %>%
  mutate(
    scram_label = ifelse(output_old_scram =="LoF", 1, 0)
  )

# === Calcolo ROC per ciascun predittore ===

roc_scram <- roc(df_pred$scram_label, -df_pred$mean_TP63_sign,  direction = "<")
roc_mutpred <- roc(df_pred$scram_label, df_pred$mutpred, direction = "<")
roc_provean <- roc(df_pred$scram_label, df_pred$provean, direction = ">")
roc_sift    <- roc(df_pred$scram_label, as.numeric(df_pred$sift),    direction = ">")
roc_polyphen<- roc(df_pred$scram_label, as.numeric(df_pred$polyphen),direction = "<")
roc_alpha   <- roc(df_pred$scram_label, df_pred$alpha,   direction = "<")
roc_cadd    <- roc(df_pred$scram_label, as.numeric(df_pred$cadd),    direction = "<")
roc_esm1b   <- roc(df_pred$scram_label, df_pred$ESM1b,   direction = ">")
roc_eve     <- roc(df_pred$scram_label, df_pred$EVE,     direction = "<")

# === Confronto visivo ROC ===

sapply(list(
  SCRAM=roc_scram, MutPred=roc_mutpred, Provean=roc_provean,
  SIFT=roc_sift, PolyPhen2=roc_polyphen, Alpha=roc_alpha,
  CADD=roc_cadd, ESM1b=roc_esm1b, EVE=roc_eve
), auc)

ggroc(list(
  "MITE" = roc_mite,
  "SCRAM TP63" = roc_scram,
  "MutPred" = roc_mutpred,
  "Provean" = roc_provean,
  "SIFT" = roc_sift,
  "PolyPhen2" = roc_polyphen,
  "PAPI" = roc_papi,
  "Alpha" = roc_alpha,
  "CADD" = roc_cadd,
  "ESM1b" = roc_esm1b,
  "EVE" = roc_eve
)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
  theme_minimal(base_size = 14) +
  labs(
    title = "ROC — Confronto predittori vs MITE (riferimento)",
    x = "False Positive Rate (1 - Specificità)",
    y = "True Positive Rate (Sensibilità)",
    color = "Predictor"
  )

#print with AUC value

roc_list <- list(
"SCRAM" <- roc(df_pred$scram_label, -df_pred$mean_TP63_sign,  direction = "<"),
"MUTPRED" <- roc(df_pred$scram_label, df_pred$mutpred, direction = "<"),
"PROVEAN" <- roc(df_pred$scram_label, as.numeric(df_pred$provean), direction = ">"),
"SIFT"    <- roc(df_pred$scram_label, as.numeric(df_pred$sift),    direction = ">"),
"POLYPHEN"<- roc(df_pred$scram_label, as.numeric(df_pred$polyphen),direction = "<"),
"ALPHA"   <- roc(df_pred$scram_label, as.numeric(df_pred$alpha),   direction = "<"),
"CADD"    <- roc(df_pred$scram_label, as.numeric(df_pred$cadd),    direction = "<"),
"ESM1B"   <- roc(df_pred$scram_label, df_pred$ESM1b,   direction = ">"),
"EVE"     <- roc(df_pred$scram_label, df_pred$EVE,     direction = "<")
)

names(roc_list) <- c(
   "SCRAM", "MutPred", "Provean", "SIFT",
  "PolyPhen2","Alpha", "CADD", "ESM1b", "EVE"
)
auc_values <- sapply(roc_list, auc)
names(roc_list) <- paste0(names(roc_list), " (AUC=", sprintf("%.2f", auc_values), ")")

ggroc(roc_list, legacy.axes = TRUE) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
  theme_minimal(base_size = 14) +
  labs(
    title = "ROC — Confronto predittori vs SCRAM_c SAM DBD(riferimento)",
    x = "False Positive Rate (1 - Specificità)",
    y = "True Positive Rate (Sensibilità)",
    color = "Predictor (AUC)"
  ) +
  theme(legend.position = "right")

ggsave("figures/ROC_SCRAMc_DBD_SAM_allpredictors_avg.pdf", height=5, width=8)
```


Fig. 6A and 6B UMAP SCRAM SAM non converted
```{r}
b_clean=read.table("merged_MGI_Novaseq/sum_counts/seurat5project/data/mutations_sam_to_conversion.txt", header = T)

b=b[!is.na(b$mut_name),]
b_new <- b %>%
  extract(
    col   = mut_name,
    into  = c("prefix","position","suffix"),
    regex = "^([A-Za-z])(\\d{3})([A-Za-z])$"
  )

b_new <- b_new %>%
  mutate(
    position      = as.integer(position) + 43,
    mut_name_new = paste0(prefix, position, suffix)
  )

int_mut=unlist(strsplit("G293H - G310Q - R304Q - R279Q - R304T - R279S - R311I - G297L - T527I - H556A - L514F - C522G - L514D - C522D - L531E - L531R - L548N - I550D", " - "))
points=b_new[b_new$mut_name_new%in%int_mut,]


ggplot(coord, aes(x=umap_1, y=umap_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points, aes(fill=conversion), size=5, pch=21) + geom_text(data=points, aes(label=mut_name_new),vjust=0, hjust=0) + 
  scale_fill_brewer(palette = "Dark2")
ggsave("figures/P63_UMAP_validated_mutants.pdf", height=5, width=8)


scram_red=readRDS("merged_MGI_Novaseq/sum_counts/seurat5project/data/scram_tp63_nocc_red_pca")
DimPlot(scram_red)
dataframe=data.frame(scram_red@reductions$umap@cell.embeddings,cluster=scram_red@meta.data$integrated_snn_res.0.2)
dataframe$cells=rownames(dataframe)
print(ggplot(data = dataframe, aes_string(x = "umap_1", y = "umap_2",color="cluster")) +
  geom_point(size = 0.5) + theme_classic() + 
 theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_text(size=15),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        title = element_text(size=24)) +
  xlab("UMAP1") + 
  ylab("UMAP2"))
ggsave("figures/MYOD_UMAP_clusters_res02.pdf", height=5, width=8)

```

Fig. S6C Heatmap score SCRAM SAM non converted
```{r}
domain="SAM"

b_clean=read.table("merged_MGI_Novaseq/sum_counts/seurat5project/data/mutations_p63_to_conversion.txt", header = T)

#score MITE with all variants
sam_scores=readRDS("Results/SAM_final_score.rds")
sam_scores=sam_scores[rownames(sam_scores)!="WT",,drop=F]
rownames(sam_scores)=paste0(substr(rownames(sam_scores),1,1),as.numeric(substr(rownames(sam_scores),2,4))+43,substr(rownames(sam_scores),5,5))
sam_scores$mut_name=rownames(sam_scores)

new_mite=b_clean
new_mite <- new_mite %>%
  extract(
    col   = mut_name,
    into  = c("prefix","position","suffix"),
    regex = "^([A-Za-z])(\\d{3})([A-Za-z])$"
  )

new_mite <- new_mite %>%
  mutate(
    position      = as.integer(position) + 43,
    mut_name_new = paste0(prefix, position, suffix)
  )

new_mite$prefix=NULL
new_mite$position=NULL
new_mite$suffix=NULL
new_mite=new_mite %>% mutate(mut_name=mut_name_new)


fin_sam<- sam_scores %>%
  inner_join(
    new_mite %>% dplyr::select(mut_name,BC,Sign_test_11),
    by = "mut_name"
  )

df_sam= fin_sam %>%
  group_by(mut_name,Sign_test_11) %>%
  summarise(Sign_test_11=mean(Sign_test_11, na.rm=TRUE), .groups = "drop") 

df_sam$residues=substr(df_sam$mut_name, 1, 4)
df_sam$aa=substr(df_sam$mut_name, 5, 5)
df_sam$order=substr(df_sam$residues, 2,4)
df_sam=df_sam[order(df_sam$order, decreasing=T),]
colnames(df_sam)[1]="score"
df_sam$residues=factor(df_sam$residues, levels = rev(unique(df_sam$residues)))
sort=sort=read.table("Results/order_aa_heatmaps.txt", header = T)
sort=as.character(sort$aa)
df_sam$aa=factor(df_sam$aa, levels = rev(sort))

z <- as.data.frame(complete(df_sam, residues, aa))


cairo_pdf(paste0("figures/Heatmap_scores_SCRAM_nc_",domain,"newSign.pdf"), height =4,width = 16)

ggplot(z, aes(aa, residues)) + theme_void()+
  geom_tile(aes(fill = Sign_test_11),size=1, color = "white") +
  scale_fill_gradientn(colors=colorRampPalette(rev(c("darkblue", 'blue', 'white', "red", "red2")))(100)
                       , values=rescale(c(min(z$Sign_test_11[!is.na(z$Sign_test_11)]),0,max(z$Sign_test_11[!is.na(z$Sign_test_11)]))), na.value = "grey50") +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size=15,family="Courier",angle = 90, vjust = 0.5, hjust=1),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(.5,"cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size=15,family="Courier")) +
  coord_flip()
dev.off()


```

Fig. 6A UMAP SCRAM SAM non converted feature plot
```{r}
scram_sam=readRDS("merged_MGI_Novaseq/sum_counts/seurat5project/data/FLE_sam_obj_FC")

df <- FetchData(scram_sam, vars = c("umap_1", "umap_2", "Sign_test_11"))

df$group <- ifelse(df$Sign_test_11 >= 0, "High TP63Sign", "Low TP63Sign")
FeaturePlot(scram_sam,features = c("TP63","ITGB4","ITGA6","IDI2","KRT14","KRT5","KRT17","FAT2"))

gene <- c("TP63","ITGB4","ITGA6","IDI2","KRT14","KRT5","KRT17","FAT2")

feature_data <- as.data.frame(t(cbind(scram_sam[["RNA"]]$data.1.1[gene, ],scram_sam[["RNA"]]$data.2.2[gene, ])))
dataset <- as.data.frame(Embeddings(scram_sam, reduction = "umap"))

# Create a ggplot from the feature data
merged=cbind(feature_data,dataset)
pdf("figures/TP63_SAM_FeaturePlots_GeneExpression.pdf", width = 12, height = 6)
for(i in gene){
gene_col_name <- paste0(i)

print(ggplot(data = merged, aes(x = umap_1, y = umap_2, color = .data[[gene_col_name]])) +
  geom_point(size = 0.9) + 
  theme_classic() + 
  labs(title = paste("Feature Plot of", gene_col_name)) + 
  xlab("UMAP1") + 
  ylab("UMAP2") + 
  scale_color_gradient(low = "grey90", high = "#6959CD") +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_text(size = 15),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5),
        title = element_text(size = 12)))

}
dev.off()

```

Fig. 6B UMAP SCRAM SAM non converted density umap
```{r}
scram_sam=readRDS("merged_MGI_Novaseq/sum_counts/seurat5project/data/FLE_sam_obj_FC")

df <- FetchData(scram_myod, vars = c("umap_1", "umap_2", "Sign_test_11"))

df$group <- ifelse(df$Sign_test_11 >= 0, "High TP63Sign", "Low TP63Sign")


p=ggplot() +
  geom_point(data = df, aes(umap_1, umap_2), color = "grey90", size = 0.3) +
  stat_density_2d_filled(
    data = df %>% filter(group == "Low TP63Sign"),
    aes(x = umap_1, y = umap_2, fill = after_stat(level)),
    geom = "polygon", color = "white"
  ) +
 scale_fill_viridis_d(option = "inferno", name = "Density") +
  coord_cartesian(
    xlim = c(min(df$umap_1) - 1, max(df$umap_1) + 1),
    ylim = c(min(df$umap_2) - 1, max(df$umap_2) + 1)
  ) +
  theme_bw() 
p2=ggplot() +
  geom_point(data = df, aes(umap_1, umap_2), color = "grey90", size = 0.3) +
  stat_density_2d_filled(
    data = df %>% filter(group == "High TP63Sign"),
    aes(x = umap_1, y = umap_2, fill = after_stat(level)),
    geom = "polygon", color = "white"
  ) +
 scale_fill_viridis_d(option = "inferno", name = "Density") +
  coord_cartesian(
    xlim = c(min(df$umap_1) - 1, max(df$umap_1) + 1),
    ylim = c(min(df$umap_2) - 1, max(df$umap_2) + 1)
  ) +
  theme_bw() 
  
pdf("figures/density_umap_tp63_2.pdf", height = 8, width = 12)
p2
dev.off()
```

Fig 6C UMAP SCRAM SAM non converted neutral and LoF mutations pinpointed
```{r}
b_clean=read.table("merged_MGI_Novaseq/sum_counts/seurat5project/data/mutations_sam_to_conversion.txt", header = T)
new_mite=b_clean

points_df=new_mite[!(is.na(new_mite$mut_name)),]
points_df$output_scram=ifelse(points_df$Sign_test_11>= 0, "Neutral mutation","LoF mutation")
points_lof_scram=points_df[points_df$output_scram %in%"LoF mutation",]
points_neutral_scram=points_df[points_df$output_scram%in%"Neutral mutation",]


ggplot(coord, aes(x=umap_1, y=umap_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_lof_scram,fill="tomato",size=5, pch=21) +  ggtitle("P63 LoF based on SCRAM")
ggsave("figures/TP63_UMAP_lof_mutants_SCRAM.pdf", height=5, width=8)

ggplot(coord, aes(x=umap_1, y=umap_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_neutral_scram,fill="forestgreen", size=5, pch=21) +  ggtitle("P63 neutral based on SCRAM")
ggsave("figures/TP63_UMAP_neutral_mutants_SCRAM.pdf", height=5, width=8)


points_df=write.csv(points_df,"merged_MGI_Novaseq/sum_counts/seurat5project/data/sam_list_variants_neutral_lof.csv")

points_df=new_mite[!(is.na(new_mite$mut_name)),]
feature_data <- as.data.frame(t(cbind(scram_sam[["RNA"]]$data.1.1[gene, ],scram_sam[["RNA"]]$data.2.2[gene, ])))
metadata=scram_sam@meta.data
coord=merge(sam_scram@meta.data, sam_scram@reductions$umap@cell.embeddings, by=0)
coord$BC=coord$id
coord$ITGB4=feature_data$ITGB4
coord$test_itgb4=ifelse(coord$ITGB4>0, "Neutral mutation","LoF mutation")

fin_coord <- points_df %>%
  inner_join(
    coord %>% dplyr::select(BC, ITGB4, test_itgb4),
    by = "BC"
  )


points_lof_itgb4=fin_coord[fin_coord$test_itgb4 %in%"LoF mutation",]
points_neutral_itgb4=fin_coord[fin_coord$test_itgb4%in%"Neutral mutation",]


ggplot(coord, aes(x=umap_1, y=umap_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_lof_itgb4,fill="tomato",size=5, pch=21) +  ggtitle("P63 LoF based on ITGB4")
ggsave("figures/P63_UMAP_ITGB4_lof_mutants_SCRAM.pdf", height=5, width=8)

ggplot(coord, aes(x=umap_1, y=umap_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_neutral_itgb4,fill="forestgreen", size=5, pch=21) +  ggtitle("P63 neutral based on ITGB4")
ggsave("figures/P63_UMAP_ITGB4_neutral_mutants_SCRAM.pdf", height=5, width=8)

```

Fig 6D  mean score SCRAM SAM non converted
```{r}
domain="SAM"

sam_scores=readRDS("Results/SAM_final_score.rds")
sam_scores=sam_scores[rownames(sam_scores)!="WT",,drop=F]
rownames(sam_scores)=paste0(substr(rownames(sam_scores),1,1),as.numeric(substr(rownames(sam_scores),2,4))+43,substr(rownames(sam_scores),5,5))
sam_scores$mut_name=rownames(sam_scores)

new_mite=b_clean
new_mite <- new_mite %>%
  extract(
    col   = mut_name,
    into  = c("prefix","position","suffix"),
    regex = "^([A-Za-z])(\\d{3})([A-Za-z])$")

new_mite <- new_mite %>%
  mutate(
    position      = as.integer(position) + 43,
    mut_name_new = paste0(prefix, position, suffix))


new_mite$prefix=NULL
new_mite$position=NULL
new_mite$suffix=NULL
new_mite=new_mite %>% mutate(mut_name=mut_name_new)

fin_sam<- sam_scores %>%
  inner_join(
    new_mite %>% dplyr::select(mut_name,BC,Sign_test_11),
    by = "mut_name")

df_sam= fin_sam %>%
  group_by(mut_name,Sign_test_11) %>%
  summarise(Sign_test_11=mean(Sign_test_11, na.rm=TRUE), .groups = "drop") 

df_sam$residues=substr(df_sam$mut_name, 1, 4)
df_sam$aa=substr(df_sam$mut_name, 5, 5)
df_sam$order=substr(df_sam$residues, 2,4)
df_sam=df_sam[order(df_sam$order, decreasing=T),]
colnames(df_sam)[1]="score"
df_sam$residues=factor(df_sam$residues, levels = rev(unique(df_sam$residues)))
sort=sort=read.table("../Paper_FP/Results//order_aa_heatmaps.txt", header = T)
sort=as.character(sort$aa)
df_sam$aa=factor(df_sam$aa, levels = rev(sort))

sub=df_sam

sub=sub[order(sub$order, decreasing=T),]
sub$residues=as.character(sub$residues)
sort=sort=read.table("../Paper_FP/Results//order_aa_heatmaps.txt", header = T)
sort=as.character(sort$aa)
sub$aa=factor(sub$aa, levels = rev(sort))
sub$residues=factor(sub$residues, levels = unique(sub$residues))

plot <- data.frame(
  residues = names(tapply(sub$Sign_test_11, sub$residues, mean, na.rm = TRUE)),
  mean_score = as.numeric(tapply(sub$Sign_test_11, sub$residues, mean, na.rm = TRUE)))


plot$residues=factor(plot$residues, levels = rev(unique(plot$residues)))

pdf(paste0("figures/SCRAM_mean_newscore","_",domain,".pdf"), height =4,width = 16)
ggplot(plot, aes(residues, "Mean")) + theme_void()+
  geom_tile(aes(fill = mean_score),size=1, color = "white") +
  scale_fill_gradientn(colors=colorRampPalette(c("red2", 'red', 'white', "darkblue", "darkblue"))(100),limits = c(-1.75, 2.50), na.value = "grey50") +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size=15,family="Courier",angle = 90, vjust = 0.5, hjust=1),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(.5,"cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size=15,family="Courier"))
dev.off()
```

Fig 6E and 6H Roc curves SCRAM SAM non converted
```{r}

common_df=read.csv("merged_MGI_Novaseq/sum_counts/seurat5project/data/SAMcomparison_old_new_scram.csv")

df_summary <- common_df %>%
  group_by(mut_name, Output) %>%  # Neutral)
  summarise(
    mean_TP63 = mean(TP63_sign1, na.rm = TRUE),
    mean_Sign = mean(Sign_test_11, na.rm = TRUE),
    occurrence = n(),             
    .groups = "drop"
  )

df_summary$output_mite=df_summary$Output
df_summary$output_old_scram=ifelse(df_summary$mean_TP63>=0,"Neutral","LoF")
df_summary$output_new_scram=ifelse(df_summary$mean_Sign>=0,"Neutral","LoF")


scram_c_nc_mite <- df_summary %>% dplyr::select(mean_Sign,mut_name,mean_TP63,output_new_scram, output_old_scram,output_mite)


df_pred <- scram_c_nc_mite %>%
  mutate(
    mite_label = ifelse(output_mite == "LoF mutation", 1, 0),)

roc_list <- list(
 "MITE"  <- roc(df_pred$mite_label, df_pred$mite_label,      direction = "<"),
"SCRAM_c" <- roc(df_pred$mite_label, -df_pred$mean_TP63,  direction = "<"),
"SCRAM_nc" <- roc(df_pred$mite_label, -df_pred$mean_Sign, direction = "<")
)

names(roc_list) <- c(
  "MITE", "SCRAM_c","SCRAM_nc")

auc_values <- sapply(roc_list, auc)
names(roc_list) <- paste0(names(roc_list), " (AUC=", sprintf("%.2f", auc_values), ")")

# Plot con AUC in legenda
ggroc(roc_list, legacy.axes = TRUE) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
  theme_minimal(base_size = 14) +
  labs(
    title = "ROC — Confronto SCRAM conversion/non conversion vs MITE (riferimento)",
    x = "False Positive Rate (1 - Specificità)",
    y = "True Positive Rate (Sensibilità)",
    color = "Predictor (AUC)"
  ) +
  theme(legend.position = "right")


ggsave("figures/ROC_MITE_ncSCRAM_cSCRAM_avg.pdf", height=5, width=8)

b_clean=read.table("merged_MGI_Novaseq/sum_counts/seurat5project/data/mutations_p63_to_conversion.txt", header = T)

b_temp <- b_clean %>%
  extract(
    col   = mut_name,
    into  = c("prefix","position","suffix"),
    regex = "^([A-Za-z])(\\d{3})([A-Za-z])$"
  )

b_temp <- b_temp %>%
  mutate(
    position      = as.integer(position) + 43,
    mut_name_new = paste0(prefix, position, suffix)
  )

b_temp$prefix=NULL
b_temp$position=NULL
b_temp$suffix=NULL

b_temp$mut_name=b_temp$mut_name_new
common_df=b_temp

df_summary <- common_df %>%
  group_by(mut_name) %>%  
  summarise(
    mean_Sign = mean(Sign_test_11, na.rm = TRUE),
    occurrence = n(),             
    .groups = "drop"
  )

file_path <- "Predictors_benchmark_RDS/Output MITE and predictors.xlsx"  

sam <- read_excel(file_path, sheet = 2)
sam$mut_name=sam$Variant

scram_allpre_mite<- sam %>%
  inner_join(
    df_summary %>% dplyr::select(mut_name, mean_Sign),
    by = "mut_name"
  )

scram_allpre_mite$output_new_scram=ifelse(scram_allpre_mite$mean_Sign>=0,"Neutral","LoF")

df_pred <- scram_allpre_mite %>%
  mutate(
    scram_label = ifelse(output_new_scram =="LoF", 1, 0)
  )


roc_scram <- roc(df_pred$scram_label, -df_pred$mean_Sign,  direction = "<")
roc_mutpred <- roc(df_pred$scram_label, df_pred$mutpred, direction = "<")
roc_provean <- roc(df_pred$scram_label, df_pred$provean, direction = ">")
roc_sift    <- roc(df_pred$scram_label, df_pred$sift,    direction = ">")
roc_polyphen<- roc(df_pred$scram_label, df_pred$polyphen,direction = "<")
roc_papi    <- roc(df_pred$scram_label, as.numeric(df_pred$papi),    direction = "<")
roc_alpha   <- roc(df_pred$scram_label, df_pred$alpha,   direction = "<")
roc_cadd    <- roc(df_pred$scram_label, df_pred$cadd,    direction = "<")
roc_esm1b   <- roc(df_pred$scram_label, df_pred$ESM1b,   direction = ">")
roc_eve     <- roc(df_pred$scram_label, df_pred$EVE,     direction = "<")


sapply(list(
  SCRAM=roc_scram, MutPred=roc_mutpred, Provean=roc_provean,
  SIFT=roc_sift, PolyPhen2=roc_polyphen, PAPI=roc_papi, Alpha=roc_alpha,
  CADD=roc_cadd, ESM1b=roc_esm1b, EVE=roc_eve
), auc)

ggroc(list(
  "MITE" = roc_mite,
  "SCRAM Sign" = roc_scram,
  "MutPred" = roc_mutpred,
  "Provean" = roc_provean,
  "SIFT" = roc_sift,
  "PolyPhen2" = roc_polyphen,
  "PAPI" = roc_papi,
  "Alpha" = roc_alpha,
  "CADD" = roc_cadd,
  "ESM1b" = roc_esm1b,
  "EVE" = roc_eve
)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
  theme_minimal(base_size = 14) +
  labs(
    title = "ROC — Confronto predittori vs MITE (riferimento)",
    x = "False Positive Rate (1 - Specificità)",
    y = "True Positive Rate (Sensibilità)",
    color = "Predictor"
  )

#print with AUC value

roc_list <- list(
"SCRAM" <- roc(df_pred$scram_label, -df_pred$mean_Sign,  direction = "<"),
"MUTPRED" <- roc(df_pred$scram_label, df_pred$mutpred, direction = "<"),
"PROVEAN" <- roc(df_pred$scram_label, df_pred$provean, direction = ">"),
"SIFT"    <- roc(df_pred$scram_label, df_pred$sift,    direction = ">"),
"POLYPHEN"<- roc(df_pred$scram_label, df_pred$polyphen,direction = "<"),
"ALPHA"   <- roc(df_pred$scram_label, df_pred$alpha,   direction = "<"),
"CADD"    <- roc(df_pred$scram_label, df_pred$cadd,    direction = "<"),
"ESM1B"   <- roc(df_pred$scram_label, df_pred$ESM1b,   direction = ">"),
"EVE"     <- roc(df_pred$scram_label, df_pred$EVE,     direction = "<")
)

names(roc_list) <- c(
   "SCRAM", "MutPred", "Provean", "SIFT",
  "PolyPhen2","Alpha", "CADD", "ESM1b", "EVE"
)
auc_values <- sapply(roc_list, auc)
names(roc_list) <- paste0(names(roc_list), " (AUC=", sprintf("%.2f", auc_values), ")")

# Plot con AUC in legenda
ggroc(roc_list, legacy.axes = TRUE) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
  theme_minimal(base_size = 14) +
  labs(
    title = "ROC — Confronto predittori vs SCRAM_nc (riferimento)",
    x = "False Positive Rate (1 - Specificità)",
    y = "True Positive Rate (Sensibilità)",
    color = "Predictor (AUC)"
  ) +
  theme(legend.position = "right")

ggsave("figures/ROC_SCRAMnc_allpredictors_avg_newsign.pdf", height=5, width=8)
```

Fig 6I and 6J violin plot Neutral and LoF variants SCRAM SAM non converted
```{r}
b_clean=read.table("merged_MGI_Novaseq/sum_counts/seurat5project/data/mutations_p63_to_conversion.txt", header = T)

b_clean$output_scram=ifelse(b_clean$Sign_test_11>= 0, "Neutral","LoF")

df <- b_clean %>%
  dplyr::select(BC, mut_name, Sign_test_11, output_scram) %>%
  drop_na(BC, mut_name, Sign_test_11, output_scram)

cell_counts <- df %>%
  group_by(mut_name, output_scram) %>%   
  summarise(occurrence = n(), .groups = "drop") %>%
  arrange(desc(occurrence))

top5_neutral <- cell_counts %>%
  filter(output_scram == "Neutral") %>%
  top_n(5,occurrence ) %>%
  pull(mut_name)
top5_neutral=top5_neutral[1:5]

top5_lof <- cell_counts %>%
  filter(output_scram == "LoF") %>%
  top_n(5, occurrence) %>%
  pull(mut_name)
top5_lof=top5_lof[1:5]
selected_variants <- unique(c(top5_neutral, top5_lof))

df_subset <- df %>%
  filter(mut_name %in% selected_variants)

df_subset$mut_name <- factor(df_subset$mut_name,
                              levels = selected_variants)


#adding statistical test Mann-Whitney on all the variants

df$score=df$Sign_test_11
test_stat_all <- wilcox.test(score ~ output_scram, data = df)
my_comparisons <- list(c("Neutral", "LoF"))

variant_counts <- b_clean %>%
  distinct(mut_name, output_scram) %>% 
  group_by(output_scram) %>%
  summarise(N_variants = n())

labels_df <- data.frame(
  output_scram = variant_counts$output_scram,
  label_text = paste0(variant_counts$output_scram, "\n(N=", variant_counts$N_variants, ")")
)

x_labels <- setNames(labels_df$label_text, labels_df$output_scram)
ggplot(df, aes(x = output_scram, y = Sign_test_11, fill = output_scram)) +
  geom_violin(trim = FALSE, scale = "width") +
  geom_boxplot(width = 0.1, outlier.shape = NA, color = "black") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Signature distribution by variant (SCRAM-seq)",
    x = "Classification Group",
    y = "TP63 signature score"
  ) +
  scale_fill_manual(values = c("Neutral" = "#4daf4a", "LoF" = "#e41a1c")) +
  scale_x_discrete(labels = x_labels)+
  stat_compare_means(
    comparisons = my_comparisons,
    method = "wilcox.test", 
    label = "p.signif",  
    hide.ns = TRUE,      
    vjust = 0.5,            
    label.y.npc = "top"   
  ) +
  theme(
    axis.text.x = element_text(angle = 0), 
    legend.position = "top"
  )
ggsave("figures/violin_plot_significance_all_variants_newsignature_SAM.pdf", width = 9, height = 6)



df <- b_clean %>%
  dplyr::select(BC, mut_name, Sign_test_11) %>%
  drop_na() %>%
  mutate(
    Expression_Status = ifelse(Sign_test_11 >= 0, "Neutral", "LoF"))

cell_summary <- df %>%
  group_by(Expression_Status) %>%
  summarise(n_cells = n(), .groups = "drop")

x_labels <- setNames(
  paste0(cell_summary$Expression_Status, "\n(n = ", cell_summary$n_cells, " cells)"),
  cell_summary$Expression_Status
)
my_comparisons <- list(c("Neutral", "LoF"))

ggplot(df, aes(x = Expression_Status, y = Sign_test_11, fill = Expression_Status)) +
  geom_violin(trim = FALSE, scale = "width") +
  geom_boxplot(width = 0.1, outlier.shape = NA, color = "black") +
  stat_compare_means(
    comparisons = my_comparisons,
    method = "wilcox.test",
    label = "p.format",    
    hide.ns = TRUE,
    vjust = 0.5,           
    label.y.npc = "top" ,
    size = 5,
    fontface = "bold"
  ) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Signature Expression Profile (n = 1216 cells)",
    subtitle = "Distribution of all sequenced cells based on transcriptional response",
    x = "Transcriptional State",
    y = "TP63 Signature Score",
    fill = "Status"
  ) +
  scale_fill_manual(values = c("Neutral" = "#4daf4a", "LoF" = "#e41a1c")) +
  scale_x_discrete(labels = x_labels)+
  theme(
    axis.text.x = element_text(angle = 0), 
    legend.position = "top"
  ) 

ggsave("figures/violin_SAMnc_global_cell_distribution_signature.pdf", width = 8, height = 7)


df_subset$score=df_subset$Sign_test_11
test_stat=wilcox.test(score ~ output_scram, data = df_subset)

p_value <- test_stat$p.value 
p_value_label <- ""

if (p_value <= 0.0001) {
    p_value_label <- "****"
} else if (p_value <= 0.001) {
    p_value_label <- "***"
} else if (p_value <= 0.01) {
    p_value_label <- "**"
} else if (p_value <= 0.05) {
    p_value_label <- "*"
} else {
    p_value_label <- "ns" 
}


 ggplot(df_subset, aes(x = mut_name, y = Sign_test_11 , fill = output_scram)) +
  geom_violin(trim = FALSE, scale = "width") +
  geom_boxplot(width = 0.1, outlier.shape = NA, color = "black") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Signature distribution by variant (SCRAM-seq)",
    x = "Variant",
    y = "TP63 signature score"
  ) +
  scale_fill_manual(values = c("Neutral" = "#4daf4a", "LoF" = "#e41a1c")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  ) +
  geom_segment(
    aes(x = 1, xend = 5, y = y_max, yend = y_max),
    color = "black", linewidth = 0.5
  ) +
  geom_segment(
    aes(x = 6, xend = 10, y = y_max, yend = y_max),
    color = "black", linewidth = 0.5
  ) +
    geom_segment(
    aes(x = 1, xend = 1, y = y_max, yend = y_max - 0.05),
    color = "black", linewidth = 0.5
  ) +
  geom_segment(
    aes(x = 10, xend = 10, y = y_max, yend = y_max - 0.05),
    color = "black", linewidth = 0.5
  ) +
annotate(
    "text",
    x = 5.5, 
    y = y_pvalue,
    label = p_value_label, 
    size = 6, 
    fontface = "bold"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))

ggsave("figures/violin_plot_significance_top5_variants_newsignature_SAM.pdf", width = 9, height = 6)
```

Fig 7A SCRAM MYOD1 featureplots
```{r}
scram_myod=readRDS("merged_MGI_Novaseq/sum_counts/seurat5project/data/scram_myod_nocc_red_pca")
FeaturePlot(scram_myod,features =c("CHRNA1","DYSF","TNNT2","TTN","SGCA","MUSK","ACTC1","TRIM72","TPM1","ITGA7","TNNT1","ENO3"))

feature_data <- as.data.frame(t(cbind(scram_myod[["RNA"]]$data.1.1[gene2, ],scram_myod[["RNA"]]$data.2.2[gene2, ])))
dataset <- as.data.frame(Embeddings(scram_myod, reduction = "umap"))


merged=cbind(feature_data,dataset)
pdf("figures/MYOD1_FeaturePlots_GeneExpression.pdf", width = 12, height = 6)
for(i in gene2){
gene_col_name <- paste0(i)

print(ggplot(data = merged, aes(x = umap_1, y = umap_2, color = .data[[gene_col_name]])) +
  geom_point(size = 0.9) + 
  theme_classic() + 
  labs(title = paste("Feature Plot of", gene_col_name)) + 
  xlab("UMAP1") + 
  ylab("UMAP2") + 
  scale_color_gradient(low = "grey90", high = "#6959CD") +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_text(size = 15),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5),
        title = element_text(size = 12)))

}
dev.off()
```

Fig 7B SCRAM MYOD1 density UMAP
```{r}
scram_myod=readRDS("merged_MGI_Novaseq/sum_counts/seurat5project/data/FLE_myod1_obj_FC")
scram_sam=readRDS("merged_MGI_Novaseq/sum_counts/seurat5project/data/FLE_sam_obj_FC")

df <- FetchData(scram_myod, vars = c("umap_1", "umap_2", "Sign_test_11"))

df$group <- ifelse(df$Sign_test_11 >= 0, "High TP63Sign", "Low TP63Sign")


 p1 <- ggplot() +
   geom_point(data = df, aes(umap_1, umap_2), color = "grey90", size = 0.3) +
   stat_density_2d(data = df %>% filter(group == "Low MYOD1Sign"),
                   aes(x = umap_1, y = umap_2, fill = after_stat(level)),
                   geom = "polygon", color = "black",contour_var = "ndensity", n=200,  expand = 1.2) +
   scale_fill_viridis_c(option = "inferno", name = "Density") +
   coord_cartesian(xlim=c(min(df$umap_1)-1,max(df$umap_1)+1),
                   ylim=c(min(df$umap_2)-1,max(df$umap_2)+1), expand=TRUE) +
   theme_bw() +
   labs(title = "Low MYOD1Sign") +  coord_equal()
 p1
 
 p2 <- ggplot() +
   geom_point(data = df, aes(umap_1, umap_2), color = "grey90", size = 0.3) +
   stat_density_2d(data = df %>% filter(group == "High MYOD1Sign"),
                   aes(x = umap_1, y = umap_2, fill = after_stat(level)),
                   geom = "polygon", color = "black") +
   scale_fill_viridis_c(option = "inferno", name = "Density") +
   theme_bw() +
   labs(title = "High MYOD1Sign") +
   coord_equal()
 
 pdf("figures/density_umap_myod1", height = 8, width = 12)
  p1+p2
 dev.off()
```

Fig. 7C SCRAM MYOD1 UMAP neutral LoF mutations
```{r}

b=read.table("merged_MGI_Novaseq/sum_counts/seurat5project/data/myod_variants_score_uncleaned_new.txt",sep = "\t")

points=b
b_clean <- b[!is.na(b$mut_name) & b$mut_name != "", ]
b_temp=b_clean[,c(1,3,4,5,6,7,8,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28)]
b_temp$id=NULL

new_mite=b_temp

points_df=new_mite[!(is.na(new_mite$mut_name)),]
points_df$output_scram=ifelse(points_df$Sign_test_11>= 0, "Neutral mutation","LoF mutation")
points_lof_scram=points_df[points_df$output_scram %in%"LoF mutation",]
points_neutral_scram=points_df[points_df$output_scram%in%"Neutral mutation",]

coord=read.csv("merged_MGI_Novaseq/sum_counts/seurat5project/data/scram_red_myod_coord.csv")

ggplot(coord, aes(x=umap_1, y=umap_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_lof_scram,fill="tomato",size=5, pch=21) +  ggtitle("MYOD1 LoF based on SCRAM")
ggsave("figures/MYOD1_UMAP_lof_mutants_SCRAM.pdf", height=5, width=8)

ggplot(coord, aes(x=umap_1, y=umap_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=points_neutral_scram,fill="forestgreen", size=5, pch=21) +  ggtitle("MYOD1 neutral based on SCRAM")
ggsave("figures/MYOD1_UMAP_neutral_mutants_SCRAM.pdf", height=5, width=8)

point_all_variants=rbind(points_lof_scram,points_neutral_scram)

ggplot(coord, aes(x=umap_1, y=umap_2))+
geom_point(size=.5, color="grey90") + theme_bw()+theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "none") + xlab("UMAP1") + ylab("UMAP2") +  geom_point(data=point_all_variants,aes(fill=output_scram),size=5, pch=21) +  ggtitle("P63 LoF based on SCRAM")
ggsave("fffigures/MYOD1_UMAP_all_mutants_SCRAM.pdf", height=5, width=8)
```

Fig. 7D heatmap scor SCRAM MYOD1
```{r}

domain="MYOD1"
b_clean=read.table("merged_MGI_Novaseq/sum_counts/seurat5project/data/mutations_myod_to_conversion.txt", header = T)

new_mite=b_clean
fin_myod<- new_mite %>% dplyr::select(mut_name,BC,Sign_test_11)

df_myod= fin_myod %>%
  group_by(mut_name,Sign_test_11) %>%
  summarise(Sign_test_11=mean(Sign_test_11, na.rm=TRUE), .groups = "drop") 

df_myod$residues=str_match(df_myod$mut_name, "^([A-Z][0-9]+)")[,2]
df_myod$aa=str_match(df_myod$mut_name, "([A-Z])$")[,2]
df_myod$order=str_match(df_myod$mut_name, "^[A-Z]([0-9]+)")[,2]

df_myod$order=as.numeric(df_myod$order)
df_myod=df_myod[order(df_myod$order, decreasing=T),]
colnames(df_myod)[1]="score"
df_myod$residues=factor(df_myod$residues, levels = rev(unique(df_myod$residues)))

sort=sort=read.table("../Paper_FP/Results//order_aa_heatmaps.txt", header = T)
sort=as.character(sort$aa)
df_myod$aa=factor(df_myod$aa, levels = rev(sort))

z <- as.data.frame(complete(df_myod, residues, aa))

cairo_pdf(paste0("figures/Heatmap_scores_SCRAM_nc_",domain,"newsign2_rescale.pdf"), height =4,width = 16)

ggplot(z, aes(aa, residues)) + theme_void()+
  geom_tile(aes(fill = Sign_test_11),size=1, color = "white") +
  scale_fill_gradientn(colors=colorRampPalette(rev(c("darkblue", 'blue', 'white', "red", "red2")))(100), limits=c(-1,1),
                       , 
values=rescale(c(-1,0,1)), na.value = "grey50") +                       
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size=15,family="Courier",angle = 90, vjust = 0.5, hjust=1),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(.5,"cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size=15,family="Courier")) +
  coord_flip()
dev.off()
```

Fig. 7E and 7F violin plot MYOD1 LoF and neutral variants
```{r}
points_df=read.csv("merged_MGI_Novaseq/sum_counts/seurat5project/data/myod_list_variants_neutral_lof.csv")

df <- points_df %>%
  dplyr::select(BC, mut_name, Sign_test_11, output_scram) %>%
  drop_na(BC, mut_name, Sign_test_11, output_scram)

cell_counts <- df %>%
  group_by(mut_name, output_scram) %>%   
  summarise(occurrence = n(), .groups = "drop") %>%
  arrange(desc(occurrence))

unique_variants <- cell_counts %>%
  group_by(mut_name) %>%
  filter(n_distinct(output_scram) == 1) %>%   
  ungroup()

top5_neutral <- unique_variants %>%
  filter(output_scram == "Neutral mutation") %>%
  top_n(5,occurrence ) %>%
  pull(mut_name)
top5_neutral=top5_neutral[1:5]

top5_lof <- unique_variants %>%
  filter(output_scram == "LoF mutation") %>%
  top_n(5, occurrence) %>%
  pull(mut_name)
top5_lof=top5_lof[1:5]
selected_variants <- c(top5_neutral, top5_lof)

df_subset <- df %>%
  filter(mut_name %in% selected_variants)

df_subset$mut_name <- factor(df_subset$mut_name,
                              levels = selected_variants)

df_subset$output=ifelse(df_subset$output_scram=="Neutral mutation","Neutral","LoF")

mutation_counts <- df_subset %>%
  group_by(mut_name) %>%
  summarise(
    n_cells = n(),                                     
    mean_signature = mean(Sign_test_11, na.rm = TRUE), 
    status = unique(output)[1]                       
  ) %>%
  arrange(desc(n_cells))                              


df <- points_df %>%
  dplyr::select(BC, mut_name, Sign_test_11) %>%
  drop_na() %>%
  mutate(
    Expression_Status = ifelse(Sign_test_11 >= 0, "Neutral", "LoF"))

cell_summary <- df %>%
  group_by(Expression_Status) %>%
  summarise(n_cells = n(), .groups = "drop")

x_labels <- setNames(
  paste0(cell_summary$Expression_Status, "\n(n = ", cell_summary$n_cells, " cells)"),
  cell_summary$Expression_Status
)
my_comparisons <- list(c("Neutral", "LoF"))

#adding statistical test Mann-Whitney on all the variants

df$score=df$Sign_test_11
test_stat_all <- wilcox.test(score ~ output_scram, data = df)
my_comparisons <- list(c("Neutral mutation", "LoF mutation"))

variant_counts <- df %>%
  distinct(mut_name, output_scram) %>% 
  group_by(output_scram) %>%
  summarise(N_variants = n())

labels_df <- data.frame(
  output_scram = variant_counts$output_scram,
  label_text = paste0(variant_counts$output_scram, "\n(N=", variant_counts$N_variants, ")")
)

x_labels <- setNames(labels_df$label_text, labels_df$output_scram)

ggplot(df, aes(x = output_scram, y = Sign_test_11, fill = output_scram)) +
  geom_violin(trim = FALSE, scale = "width") +
  geom_boxplot(width = 0.1, outlier.shape = NA, color = "black") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Signature distribution by variant (SCRAM-seq)",
    x = "Classification Group",
    y = "TP63 signature score"
  ) +
  scale_fill_manual(values = c("Neutral mutation" = "#4daf4a", "LoF mutation" = "#e41a1c")) +
  scale_x_discrete(labels = x_labels)+
  stat_compare_means(
    comparisons = my_comparisons,
    method = "wilcox.test", 
    label = "p.signif",     
    hide.ns = TRUE,         
    vjust = 0.5,          
    label.y.npc = "top"  
  ) +
  theme(
    axis.text.x = element_text(angle = 0), 
    legend.position = "top"
  )
ggsave("figures/violin_plot_significance_all_variants_newsignature_MYOD1.pdf", width = 9, height = 6)

#statistical test only on the 5 selected variants
df_subset$score=df_subset$Sign_test_11
test_stat=wilcox.test(score ~ output_scram, data = df_subset)

p_value <- test_stat$p.value 

p_value_label <- ""

if (p_value <= 0.0001) {
    p_value_label <- "****"
} else if (p_value <= 0.001) {
    p_value_label <- "***"
} else if (p_value <= 0.01) {
    p_value_label <- "**"
} else if (p_value <= 0.05) {
    p_value_label <- "*"
} else {
    p_value_label <- "ns"
}

 ggplot(df_subset, aes(x = mut_name, y = Sign_test_11 , fill = output_scram)) +
  geom_violin(trim = FALSE, scale = "width") +
  geom_boxplot(width = 0.1, outlier.shape = NA, color = "black") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Signature distribution by variant (SCRAM-seq)",
    x = "Variant",
    y = "TP63 signature score"
  ) +
  scale_fill_manual(values = c("Neutral mutation" = "#4daf4a", "LoF mutation" = "#e41a1c")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  ) +
  geom_segment(
    aes(x = 1, xend = 5, y = y_max, yend = y_max),
    color = "black", linewidth = 0.5
  ) +
  geom_segment(
    aes(x = 6, xend = 10, y = y_max, yend = y_max),
    color = "black", linewidth = 0.5
  ) +
    geom_segment(
    aes(x = 1, xend = 1, y = y_max, yend = y_max - 0.05),
    color = "black", linewidth = 0.5
  ) +
  geom_segment(
    aes(x = 10, xend = 10, y = y_max, yend = y_max - 0.05),
    color = "black", linewidth = 0.5
  ) +
annotate(
    "text",
    x = 5.5, 
    y = y_pvalue,
    label = p_value_label, 
    size = 6, 
    fontface = "bold"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))

ggsave("figures/violin_plot_significance_top5_variants_newsignature_MYOD.pdf", width = 9, height = 6)

```


Fig. S7A featureplot UMAP SCRAM MYOD1 
```{r}
domain="MYOD1"
b_clean=read.table("merged_MGI_Novaseq/sum_counts/seurat5project/data/mutations_myod_to_conversion.txt", header = T)

new_mite=b_clean
fin_myod<- new_mite %>% dplyr::select(mut_name,BC,Sign_test_11)

df_myod= fin_myod %>%
  group_by(mut_name,Sign_test_11) %>%
  summarise(Sign_test_11=mean(Sign_test_11, na.rm=TRUE), .groups = "drop") 

df_myod$residues=str_match(df_myod$mut_name, "^([A-Z][0-9]+)")[,2]
df_myod$aa=str_match(df_myod$mut_name, "([A-Z])$")[,2]
df_myod$order=str_match(df_myod$mut_name, "^[A-Z]([0-9]+)")[,2]

df_myod$order=as.numeric(df_myod$order)
df_myod=df_myod[order(df_myod$order, decreasing=T),]
colnames(df_myod)[1]="score"
df_myod$residues=factor(df_myod$residues, levels = rev(unique(df_myod$residues)))

sort=sort=read.table("Results/order_aa_heatmaps.txt", header = T)
sort=as.character(sort$aa)
df_myod$aa=factor(df_myod$aa, levels = rev(sort))

z <- as.data.frame(complete(df_myod, residues, aa))


cairo_pdf(paste0("figures/Heatmap_scores_SCRAM_nc_",domain,"newsign2_rescale.pdf"), height =4,width = 16)

ggplot(z, aes(aa, residues)) + theme_void()+
  geom_tile(aes(fill = Sign_test_11),size=1, color = "white") +
  scale_fill_gradientn(colors=colorRampPalette(rev(c("darkblue", 'blue', 'white', "red", "red2")))(100), limits=c(-1,1),
                       , 
values=rescale(c(-1,0,1)), na.value = "grey50") +                       
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size=15,family="Courier",angle = 90, vjust = 0.5, hjust=1),
    axis.text.y = element_text(size=15,family="Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(.5,"cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size=15,family="Courier")) +
  coord_flip()
dev.off()




pred=read.table("Predictors_benchmark_RDS/output vep myod1.txt", header = T)
pred=pred[,c("Protein_position","Amino_acids","Codons","SIFT","PolyPhen")]


df <- pred %>%
  mutate(
    mut = paste0(
      stringr::str_extract(Amino_acids, "^[A-Z]"),  
      Protein_position,
      stringr::str_extract(Amino_acids, "(?<=/)[A-Z]")
    ),
    SIFT_score = as.numeric(str_extract(SIFT, "\\d+\\.\\d+")),
    PolyPhen_score = as.numeric(str_extract(PolyPhen, "\\d+\\.\\d+"))
  )

df$SIFT=replace_na(df$SIFT_score, 0)
df$Polyphen=replace_na(df$PolyPhen_score, 0)

input_myod=data.frame(mut_name=df$mut,sift=df$SIFT,polyphen=df$Polyphen)

sort=read.table("Results/order_aa_heatmaps.txt",header=T)
sort=as.character(sort$aa)

rownames(input_myod)=input_myod$mut_name
input_myod$mut_name=NULL


mut_pred=input_myod[,"sift",drop=F]
mut_pred$orig <- substr(rownames(mut_pred), 1, 1)
mut_pred$aa= substr(rownames(mut_pred), nchar(rownames(mut_pred)), nchar(rownames(mut_pred)))
mut_pred$order=substr(rownames(mut_pred), 2, nchar(rownames(mut_pred)) - 1)
mut_pred$residues=paste0(mut_pred$orig,mut_pred$order)
mut_pred$order=as.numeric(mut_pred$order)
mut_pred=mut_pred[order(mut_pred$order, decreasing=T),]
colnames(mut_pred)[1]="score"
mut_pred$residues=factor(mut_pred$residues, levels = (unique(mut_pred$residues)))


mut_pred$aa=factor(mut_pred$aa, levels = rev(sort))
z <- as.data.frame(complete(mut_pred, residues, aa))

#sift
limit=0.05
colors=colorRampPalette(c("red2", 'red2', 'white', "darkblue", "darkblue"))(100)


cairo_pdf(paste0("figures/Heatmap_scores_SIFT_",domain,".pdf"), height =4,width = 16)
     ggplot(z, aes(aa, residues)) +
      theme_void() +
      geom_tile(aes(fill = as.numeric(score)), size = 1, color = "white") +
      scale_fill_gradientn(
        colors = colors,
        values = rescale(c(min(as.numeric(z$score[!is.na(z$score)])), 0.05, max(as.numeric(z$score[!is.na(z$score)])))),
         limits = c(0, max(as.numeric(z$score[!is.na(z$score)]))),
        na.value = "grey50"
      ) +
      ylab("") +
      xlab("") +
      theme(
        axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 15, family = "Courier"),
        legend.title = element_blank(),
        legend.key.size = unit(.5, "cm"),
        legend.justification = c("right", "bottom"),
        legend.text = element_text(size = 15, family = "Courier")
      ) +
      coord_flip()
dev.off()

#Mean score sift MYOD1

domain="MYOD1"
plot <- data.frame(
  residues = names(tapply(mut_pred$score, mut_pred$residues, mean, na.rm = TRUE)),
  mean_score = as.numeric(tapply(mut_pred$score, mut_pred$residues, mean, na.rm = TRUE)))


plot$residues=factor(plot$residues, levels = (unique(plot$residues)))


pdf(paste0("figures/sift_mean_score","_",domain,".pdf"), height =4,width = 16)
  ggplot(plot, aes(residues, "sift")) + theme_void() +
  geom_tile(aes(fill = mean_score), size = 1, color = "white") +
  scale_fill_gradientn(
    colors = colors,
    values = rescale(c(0, limit, max(z$score, na.rm = TRUE))),
    limits = c(0, max(z$score, na.rm = TRUE)),
    na.value = "grey50"
  ) +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 15, family = "Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(0.5, "cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size = 15, family = "Courier")
  )
dev.off()

#polyphen

mut_pred=input_myod[,"polyphen",drop=F]
mut_pred$orig <- substr(rownames(mut_pred), 1, 1)
mut_pred$aa= substr(rownames(mut_pred), nchar(rownames(mut_pred)), nchar(rownames(mut_pred)))
mut_pred$order=substr(rownames(mut_pred), 2, nchar(rownames(mut_pred)) - 1)
mut_pred$residues=paste0(mut_pred$orig,mut_pred$order)
mut_pred$order=as.numeric(mut_pred$order)
mut_pred=mut_pred[order(mut_pred$order, decreasing=T),]
colnames(mut_pred)[1]="score"
mut_pred$residues=factor(mut_pred$residues, levels = rev(unique(mut_pred$residues)))


mut_pred$aa=factor(mut_pred$aa, levels = rev(sort))
z <- as.data.frame(complete(mut_pred, residues, aa))


limit=0.5
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)

cairo_pdf(paste0("merged_MGI_Novaseq/sum_counts/seurat5project/figures/Heatmap_scores_POLYPHEN_",domain,".pdf"), height =4,width = 16)
     ggplot(z, aes(aa, residues)) +
      theme_void() +
      geom_tile(aes(fill = as.numeric(score)), size = 1, color = "white") +
      scale_fill_gradientn(
        colors = colors,
        values = rescale(c(min(as.numeric(z$score[!is.na(z$score)])), limit, max(as.numeric(z$score[!is.na(z$score)])))),
         limits = c(0, max(as.numeric(z$score[!is.na(z$score)]))),
        na.value = "grey50"
      ) +
      ylab("") +
      xlab("") +
      theme(
        axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 15, family = "Courier"),
        legend.title = element_blank(),
        legend.key.size = unit(.5, "cm"),
        legend.justification = c("right", "bottom"),
        legend.text = element_text(size = 15, family = "Courier")
      ) +
      coord_flip()
dev.off()

#Mean score Polyphen MYOD1

plot <- data.frame(
  residues = names(tapply(mut_pred$score, mut_pred$residues, mean, na.rm = TRUE)),
  mean_score = as.numeric(tapply(mut_pred$score, mut_pred$residues, mean, na.rm = TRUE)))

plot$residues=factor(plot$residues, levels = (unique(plot$residues)))

limit=0.5
colors=colorRampPalette(rev(c("red2", 'red2', 'white', "darkblue", "darkblue")))(100)


pdf(paste0("figures/polyphen_mean_score","_",domain,".pdf"), height =4,width = 16)
  ggplot(plot, aes(residues, "polyphen")) + theme_void() +
  geom_tile(aes(fill = mean_score), size = 1, color = "white") +
  scale_fill_gradientn(
    colors = colors,
    values = rescale(c(0, limit, max(z$score, na.rm = TRUE))),
    limits = c(0, max(z$score, na.rm = TRUE)),
    na.value = "grey50"
  ) +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 15, family = "Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(0.5, "cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size = 15, family = "Courier")
  )
dev.off()

#esm1b

esm1b=read.csv("Predictors_benchmark_RDS/MYOD1_normalized.csv", header = T, sep=",")
esm1b=esm1b[,-(3)]
esm1b_filtered=esm1b %>%
  filter(pos >= 1 & pos <= 71)


esm1b_long=esm1b_filtered %>%
  pivot_longer(
    cols = A:Y, 
    names_to = "mut_aa", 
    values_to = "score_esm1b" 
  ) %>%
  transmute(
    Protein_position = pos,
    wt_aa=wt_aa,
    mut_aa=mut_aa,
    residues=paste0(wt_aa,Protein_position),
    score.esm1b = score_esm1b
  ) %>%
  filter(wt_aa != mut_aa) %>%
  arrange(Protein_position)

esm1b_long$wt_aa=factor(esm1b_long$wt_aa, levels = rev(unique(esm1b_long$wt_aa)))


esm1b_long$residues=factor(esm1b_long$residues, levels = (unique(esm1b_long$residues)))
esm1b_long$wt_aa=factor(esm1b_long$wt_aa, levels = rev(sort))

z <- as.data.frame(complete(esm1b_long, residues, mut_aa))


limit=-7.5
colors=colorRampPalette(rev(c("darkblue", "darkblue", 'white', "red2", 'red2')))(100)

cairo_pdf(paste0("figures/Heatmap_scores_ESM1b_",domain,".pdf"), height =4,width = 16)
      ggplot(z, aes(mut_aa, residues)) +
      theme_void() +
      geom_tile(aes(fill = as.numeric(score.esm1b)), size = 1, color = "white") +
      scale_fill_gradientn(
        colors = colors,
          values = rescale(c(min(as.numeric(z$score.esm1b[!is.na(z$score.esm1b)])), limit, max(as.numeric(z$score.esm1b[!is.na(z$score.esm1b)])))),
        limits = c(min(as.numeric(z$score.esm1b[!is.na(z$score.esm1b)])), max(as.numeric(z$score.esm1b[!is.na(z$score.esm1b)]))),  labels = scales::label_number(accuracy = 1)
      ) +
      ylab("") +
      xlab("") +
      theme(
        axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 15, family = "Courier"),
        legend.title = element_blank(),
        legend.key.size = unit(.5, "cm"),
        legend.justification = c("right", "bottom"),
        legend.text = element_text(size = 18, family = "Courier")
      ) +
      coord_flip()
dev.off()


#Mean score esm1b MYOD1

plot <- data.frame(
  residues = names(tapply(esm1b_long$score.esm1b, esm1b_long$residues, mean, na.rm = TRUE)),
  mean_score = as.numeric(tapply(esm1b_long$score.esm1b, esm1b_long$residues, mean, na.rm = TRUE)))

plot$residues=factor(plot$residues, levels = (unique(plot$residues)))


pdf(paste0("igures/esm1b_mean_score","_",domain,".pdf"), height =4,width = 16)
  ggplot(plot, aes(residues, "esm1b")) + theme_void() +
  geom_tile(aes(fill = mean_score), size = 1, color = "white") +
  scale_fill_gradientn(
    colors = colors,
 values = rescale(c(min(as.numeric(z$score.esm1b[!is.na(z$score.esm1b)])), limit, max(as.numeric(z$score.esm1b[!is.na(z$score.esm1b)])))),
        limits = c(min(as.numeric(z$score.esm1b[!is.na(z$score.esm1b)])), max(as.numeric(z$score.esm1b[!is.na(z$score.esm1b)]))),  labels = scales::label_number(accuracy = 1),
    na.value = "grey50"
  ) +
  ylab("") +
  xlab("") +
  theme(
    axis.text.x = element_text(size = 15, family = "Courier", angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 15, family = "Courier"),
    legend.title = element_blank(),
    legend.key.size = unit(0.5, "cm"),
    legend.justification = c("right", "bottom"),
    legend.text = element_text(size = 15, family = "Courier")
  )
dev.off()

```

